### 2025-05-02 Supplementary_File_Rheumatoid_Data.Rmd
added Supplementary_File_Rheumatoid_Data.Rmd, other changes to manuscript.tex

---
title: "The Rheumatoid Arthritis data  \nSupplementary File for "Detection of Epistatic Interactions in High-Dimensional Genomic Data Using Random Forests"
author: "Hawlader Al-Mamun,  Rob Dunne"
date: "Wednesday, September 25, 2024"
output: 
    bookdown::pdf_document2:
     toc: true
     toc_depth: 3
     fig_caption: yes
     citation_package: biblatex
bibliography: 
header-includes:
  \usepackage{dcolumn}
  \usepackage{placeins}
  \usepackage{float}
  \usepackage{biblatex}
  \usepackage{caption}
  \usepackage{subcaption}
---

We work through the analysis of Rheumatoid Arthritis data presented in "Detection of Epistatic Interactions in High-Dimensional Genomic Data Using Random Forests"

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=6, dpi=300,echo = FALSE)
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")

```
\clearpage
# Notes

- is the data available? What about the preprocessing? Can we include this in a vignette?
- dim(dt)   4798 94485 but the paper says 94,486
- looked at variable importance. Variables that have more splits as they are important are likely to have more consecutive splits. Should we take this into account?
- I removed splits from AllInteractions where the split is on the same variable again. This does not matter for the regression method but might for other methods (Empirical Bayes)
- There is a hotspot in chromosome 6. Did we know about this? Does it get dropped when you filter on LD
- we are looking at the interaction as the product of the two SNP variables, so it takes the values $\{0,1,2,4\}$. so we are threating it as a
  regression but it does not take the value 3.  If we treat them as factors we get levels(d1:d2) "0:0" "0:1" "0:2" "1:0" "1:1" "1:2" "2:0" "2:1" "2:2"
  9 levels. Which is a more reasonable model? 
- I look at using the **empiricalzip** (Empirical Bayes zero inflated Poisson) package to model the interaction counts and set a significance
level. The agreement with the regression testing is not great.
- we have the file "Cleaned_CaseControl_RAauto_pruned200kb_.3.bim"

# load data

- 4798 samples
- 94485 SNPs

```{r, echo=FALSE, eval=FALSE}
library(lmtest)
library(ranger)
#library(RCy3)

#Noise <- 0.1
numIte  <-  1
t <- numIte
ntree <- 1000

dirOut  <-  paste0("Results")
if(!dir.exists(dirOut)){
  dir.create(dirOut)
}
#dirOut  <-  paste0("Results/pruned200K.3r2_geno_Missing_Averaged")
dt  <-  readRDS("pruned200K.3r2_geno_Missing_Averaged.rds")
pheno <- read.table("Cleaned_CaseControl_RAauto.fam",header=F)
pheno$ID <- paste0(pheno$V1,"_",pheno$V2)
all.equal(row.names(dt),pheno$ID)
dt <- data.frame(pheno=as.factor(pheno$V6),dt)
dim(dt) #[1]  4798 94485

#94,486 SNPs (4.46 billion possible pairwise interaction tests) and 4,798 individuals.

dt$pheno <- as.factor(dt$pheno)
#dt <- dt[which(dt$treatment=="short"),-2]

# quick load of results etc
dirOut  <-  paste0("Results")
load(file="Results/var_importance_gini.Rdata")
load(file=paste0(dirOut,"/All_Interactions.Rdata"))
load(file="./Results/interaction_snps_counts_fixed.Rdata")
load(file=paste0(dirOut,"/lm_results_fixed.Rdata"))

```

# Fit a Random Forest

```{r, echo=FALSE, eval=FALSE}
##Number of trees:                  1000 
#set.seed(sample(1:1000,1))
set.seed(100)
#system.time(model.rf <- ranger(dependent.variable.name="pheno",data = dt,num.trees = ntree, mtry = ncol(dt)/3, importance = 'impurity'))
#     user    system   elapsed 
#89141.264     3.601  1568.711 
#save(model.rf, file="model.rf.Rdata")

##Number of trees:                  5000 
#system.time(model.rf <- ranger(dependent.variable.name="pheno",data = dt,num.trees = 5000, mtry = ncol(dt)/3))
# user     system    elapsed 
#443201.110      8.132   7185.540 
#Number of trees:                  5000 
#Sample size:                      4798 
#Number of independent variables:  94484 
#Mtry:                             31495 
#Target node size:                 1 
#Variable importance mode:         none 
#Splitrule:                        gini 
#OOB prediction error:             33.18 % 
#save(model.rf, file="model.rf_temp2.Rdata")

##Number of trees:   1000, importance = 'impurity'
set.seed(100)
#petrichor-i1: salloc -A OD-217714 --mem=100GB --nodes=1 --ntasks-per-node=64    -J interactive -t 6:00:00 srun --pty /bin/bash -l
#etrichor-i1: sacct -j 60755241  --format="jobid,ntasks,Elapsed,TotalCPU,MaxRSS,AveVMSize,MaxVMSize"
#JobID          NTasks    Elapsed   TotalCPU     MaxRSS  AveVMSize  MaxVMSize 
#60755241                00:13:56  00:06.682                                  
#60755241.ex+        1   00:13:56  00:00.001       995K    124080K    124080K 
#60755241.0         64   00:13:53  00:06.680 104545713K 3274946112 192422568K

system.time(model.rf <- ranger(dependent.variable.name="pheno",data = dt,num.trees = ntree, mtry = ncol(dt)/3, importance = 'impurity'))
#     user    system   elapsed 
#89397.171     4.807  1591.941 
#save(model.rf, file="model.rf_temp3.Rdata")
var_importance <- model.rf$variable.importance
head(rev(sort(var_importance)))
t2 <-count_variables(model.rf)
names(t2) <-names(var_importance)

#save(var_importance , t2, file="Results/var_importance_gini.Rdata")
#load(file="Results/var_importance_gini.Rdata")

```

# variable importance

we use the pacakge RFlocalfdr to give an empirical Bayes estimate of the number of significant SNPs.
It returns 74 SNPs as significant. See figure \@ref(fig:RFlocalfdr). Note that this is just settng a significance level. These
are the 74 SNPs with maximum variable importance from the random forest fit.

```{r RFlocalfdr, eval=TRUE, echo=FALSE, fig.cap='RFlocalfdr',  fig.align = 'center', out.width = '50%'}
knitr::include_graphics("./script3/RFlocalfdr.png")

```

```{R, echo=FALSE, eval=FALSE}
library(RFlocalfdr)
vignette(package="RFlocalfdr")
#no vignettes found -- why is this?
imp<-log(var_importance)  #6275 var_importance vales =0 
plot(density(imp),xlab="log importances",main="")

cutoffs <- c(0,1,2,3)
res.con<- determine_cutoff(imp,t2 ,cutoff=cutoffs,plot=c(0,1,2,3))
plot(cutoffs,res.con[,3],pch=15,col="red",cex=1.5,ylab="max(abs(y - t1))")

cutoffs <- c(4,5,6)
res.con2 <- determine_cutoff(imp,t2 ,cutoff=cutoffs,plot=c(4,5,6))
plot(c(0,1,2,3,4,5,6),c(res.con[,3],res.con2[,3]),pch=15,col="red",cex=1.5,ylab="max(abs(y - t1))")

temp<-imp[t2 > 3]
temp <- temp - min(temp) + .Machine$double.eps
qq <- plotQ(temp,debug.flag = 1)
ppp<-run.it.importances(qq,temp,debug.flag = 0)

png("./script3/RFlocalfdr.png")
aa<-significant.genes(ppp,temp,cutoff=0.05,debug.flag=0,do.plot=TRUE,use_95_q=TRUE)
dev.off()
length(aa$probabilities) # 74

aa<-significant.genes(ppp,temp,cutoff=0.05,debug.flag=0,do.plot=TRUE,use_95_q=FALSE)
length(aa$probabilities) #  74

length(intersect(names(aa$probabilities) ,names(rev(sort(var_importance)))[1:74])) #74
#but why is the order different

var_importance <- var_importance[match(names(aa$probabilities),names(var_importance))]
#now in the same order as aa$probabilities
cbind(aa$probabilities,var_importance)

```

# look at some SNP counts

- this is the counts of how often a variable was used for a split in RF
- the varaible importance has a very high correlation with this
- There is a hotspot in chromosome 6. See figure \@ref(fig:hotspot)

```{r hotspot, eval=TRUE, echo=FALSE, fig.cap='There is a hotspot in chromosome 6',  fig.align = 'center', out.width = '50%'}
knitr::include_graphics("./figs/hotspot.png")

```

```{R, echo=FALSE, eval=FALSE}
t2[grep("X6_32471505",names(t2))] #993 
t2[grep("X6_32771829",names(t2))] # 1532
sum(t2) #[1] 295310  all splits

# how are the SNP numbers organized 
plot(as.numeric(gsub("X[1-9]*\\_([0-9]*)","\\1", names(t2))))
# it looks like they are sequential within chromosome

cols<- as.numeric((gsub("X([0-9]*)\\_[0-9]*","\\1",  names(t2))))

#png("./figs/hotspot.png")
# Set up the main plot without x-axis ticks and labels
par(mar=c(5, 4, 4, 2) + 0.1)  # Adjust margins
plot(t2, col=cols, xaxt='n', main="SNP Counts Across Genome",xlab="counts")

# Add an inset plot for chromosome 6
# Define the plotting region for the inset
par(fig=c(0.5, 0.9, 0.5, 0.9), new=TRUE, mar=c(2, 2, 1, 1))  # Adjust margins for inset

# Extract data for chromosome 6
chrom6_data <- t2[cols == 6]

# Plot the inset
plot(chrom6_data[1350:1500], col=6, main="part of Chromosome 6", xaxt='n', yaxt='n')
#dev.off()
# there is a hotspot in this region!

t2[which(t2> 800)]
#X6_32471505 X6_32771829 
#        993        1532

```

# count the interactions in actual data

```{R, echo=FALSE,eval=FALSE}
#t2 <- system.time({

  # countMatrix <- matrix(0,ncol = ncol(dt)-1,nrow=ncol(dt)-1)
  # row.names(countMatrix) <- colnames(dt)[-1]
  # colnames(countMatrix) <- colnames(dt)[-1]
  #

  for(i in 1:ntree){
    #tree2 <- getTree(model.rf2,i,T)
    tree <- treeInfo(model.rf,i)
    #tree <- treeInfo(model.n,5)

    LeftinteractionTable  <-  data.frame(matrix(0,nrow = nrow(tree),ncol=2))
    LeftinteractionTable[,1] <- tree$splitvarName

    for(l in 1:nrow(tree)){
      if(is.na(tree[l,5])==F){
        leftChildID = tree[l,2]
        leftChildIndex = which(tree$nodeID==leftChildID)
        LeftinteractionTable[l,2]=tree$splitvarName[leftChildIndex]
      }
    }
    RightinteractionTable = data.frame(matrix(0,nrow = nrow(tree),ncol=2))
    RightinteractionTable[,1]=tree$splitvarName

    for(l in 1:nrow(tree)){
      if(is.na(tree[l,5])==F){

        rightChildID = tree[l,3]
        rightChildIndex = which(tree$nodeID==rightChildID)
        RightinteractionTable[l,2]=tree$splitvarName[rightChildIndex]
      }
    }

    LeftinteractionTable=na.omit(LeftinteractionTable)
    RightinteractionTable=na.omit(RightinteractionTable)

    allInteractions = rbind(LeftinteractionTable,RightinteractionTable)
    allInteractions[,1] = as.character(allInteractions[,1])
    allInteractions[,2] = as.character(allInteractions[,2])

    if(i ==1){
      index=which(allInteractions$X1> allInteractions$X2)
      tmp=allInteractions$X2
      tmp[index]=allInteractions$X1[index]
      allInteractions$X1[index]= allInteractions$X2[index]
      allInteractions$X2= tmp
      All_Interactions=allInteractions
      #allInteractions$X1_X2=paste0(allInteractions$X1," ",allInteractions$X2)
      #countTable=table(allInteractions$X1_X2)
      #All_count_dtf=data.frame(countTable)
    }else{
      index=which(allInteractions$X1> allInteractions$X2)
      tmp=allInteractions$X2
      tmp[index]=allInteractions$X1[index]
      allInteractions$X1[index]= allInteractions$X2[index]
      allInteractions$X2= tmp
      All_Interactions=rbind(All_Interactions,allInteractions)

    }
  }

#saveRDS(All_Interactions,paste0(dirOut,"/All_Interactions.rds"))
#All_Interactions <- readRDS(paste0(dirOut,"/All_Interactions.rds"))
#save(All_Interactions,file=paste0(dirOut,"/All_Interactions.Rdata"))
#dim(All_Interactions) #[1] 294310      2

```
```{R, echo=FALSE, eval=FALSE}
# just some checks on how  All_Interactions is organized. 
head(All_Interactions)
#            X1           X2      294310      2
#1  X6_32471505  X6_32771829
#2 X21_44075267  X6_32471505
#3 X14_84754343  X4_67191986

#X6_32471505  X6_32771829 are often consecutive splits. We count occurrences

length(grep("X6_32471505", All_Interactions[grep("X6_32771829",All_Interactions$X1),2])) #0
length(grep("X6_32471505", All_Interactions[grep("X6_32771829",All_Interactions$X2),1])) #466

# and for some random selected splits
for (i in  sample(294310,100) ){
    a1<- length(grep(All_Interactions$X1[i], All_Interactions[grep(All_Interactions$X2[i],All_Interactions$X1),2])) #0
    a2<- length(grep(All_Interactions$X2[i], All_Interactions[grep(All_Interactions$X1[i],All_Interactions$X1),2])) #0
    cat(a1, a2, "\n", sep=" ")
}
#so the splits are lexigraphically ordered i.e. where "X6_32771829 X6_32471505" occurs, it is recorded as  "X6_32471505  X6_32771829" 
#never the other way around. This makes things easier to track

```

## expected numbers of consecutive splits

- if the consecutive splits were uniformly random, then variables with a lot of splits would have a higher chance of occuring together
- consider X6_32471505 and  X6_32771829. If random these would have  50 consecutive splits, they have 466

```{R, echo=FALSE, eval=FALSE}

#so how often would we expect the consecutive splits "X6_32471505 X6_32771829"
#(prob X6_32471505)*prob( X6_32771829 )* (total number of split pairs)
#splits on X6_32771829 1532 times, mtry is 31495 , total numer of splits 295310, 94484 varaibles
(31495/ 94484 ) *(993 / 31495)* (31495/ 94484 ) *(1532 / 31495)*294120
#so 50.12057 versus 466

#what about  X6_32481210 X6_32682149
#281391  X6_32481210 X6_32682149    5
t2[grep("X6_32481210",names(t2))] # 121  
t2[grep("X6_32682149",names(t2))] #  277 
(31495/ 94484 ) *(121  / 31495)* (31495/ 94484 ) *( 277  / 31495)*294120
# so 1.1 versus 5

```

## remove "self splits"

1. how often are there self-splits i.e. consecutive splits on the same variable

  - 190 are self-splits (so not many). These are all due to two transcripts "X6_32771829",  "X6_32471505" which are 1 and 2 in importance
  - we  remove these from **All_Interactions**
  - save **interaction_snps_counts_fixed.Rdata**

b

```{R, echo=FALSE, eval=FALSE}

length(which(All_Interactions$X1==All_Interactions$X2))
# of 294310 splits, 190 are self-splits

unique(All_Interactions$X1[which(All_Interactions$X1==All_Interactions$X2)])
# "X6_32771829" "X6_32471505" 2 splits these are 1 and 2 in importance

imp <- names(rev(sort(var_importance)))
grep("X6_32771829",imp)
head(match( imp, All_Interactions$X1))
#  [1]     11      1    882   2916    772   2623   2337    290   1180  61812
# [11]   3524   5620   3520   7688   9321   9917    291   1468   1074   6360

All_Interactions <- All_Interactions[-c(which(All_Interactions$X1==All_Interactions$X2)),]  #this is where we drop the self splits
dim(All_Interactions) #[1] 294120      2
tt<- table(c(All_Interactions$X1,All_Interactions$X2))
tt<-rev(sort(tt))
plot(1:74,match(imp,names(tt))[1:74])
abline(0,1)
# this proves nothing. Counting the number of times a variable is in a split is a reasonable surrogate for importance

interaction_snps <- paste0(All_Interactions$X1," ",All_Interactions$X2)
interaction_snps_counts <- table(interaction_snps)
interaction_snps_counts <- data.frame(interaction_snps_counts)
interaction_snps_counts <- interaction_snps_counts[order(interaction_snps_counts$Freq,decreasing = T),]
head(interaction_snps_counts )
#              interaction_snps Freq   #most are 1
#              interaction_snps Freq
#281122 X6_32471505 X6_32771829  466
#281468 X6_32682149 X6_32771829  144
#280884 X6_32222493 X6_32771829  113

#was                                     before we dropped the repeated splits
#281123 X6_32471505 X6_32771829  466
#281469 X6_32682149 X6_32771829  144
#281616 X6_32771829 X6_32771829  128
# see the repeated split on X6_32771829

aa<-rep(NA, 74)
for ( i in 1:74){
    aa[i] <- length(grep(imp[i],interaction_snps_counts$interaction_snps[1:640]))
    }
aa
# [1] 301 154  43  15  15  19  16  17  13   9   7   7   6   9   8   3   6   4   7
#[20]   3   5   5   4   3   4   4   2   2   2   3   3   3   3   4   3   2   3   4
#[39]   4   1   3   2   2   1   1   4   1   2   1   2   2   3   1   4   1   3   2
#[58]   1   2   3   0   1   1   2   2   3   3   1   0   1   1   1   0   0

sum(interaction_snps_counts$Freq) # 294120

#saveRDS(interaction_snps_counts,paste0(dirOut,"/interaction_snps_counts_fixed.rds"))
#interaction_snps_counts <- readRDS("./Results/interaction_snps_counts_fixed.rds")
#save(interaction_snps_counts, file="./Results/interaction_snps_counts_fixed.Rdata")
#load(file="./Results/interaction_snps_counts_fixed.Rdata")

```

# run linear models

- we consider all interaction_snps where the frequency is greater than 2. That is 640 interactions 
- we fit generalized linear models with an interaction term and record the term and the $p$-value
- we fit models treating the SNP levels as factors, we fit a models with and witout the interaction term and do a likelihood ratio test
- we have Bonferrroni, and FDR adjusted p-values for glm regression with numeric SNP and FDR adjusted p-values for LRT on factor interactions
- the resulting data frame contains

   - interaction_snps  -- the interaction under consideration e.g. "X6_32308125 X6_32535726"
   - Freq              -- the frequency of that interaction 
   - SNP1_effect       -- Estimates
   - SNP1_Pvalue       -- Pr(>|z|)
   - SNP2_effect       -- Estimates
   - SNP2_Pvalue       -- Pr(>|z|)
   - int_effect   interaction term
   - int_Pvalue
   - correlations  cor(dt[,ind1],dt[,ind2])
   - Ch2          lrtest(nested.model, s2) d1 d2 factors  Pr(>Chisq)
   - splits_SNP1   number of times this varaible is the split varaible 
   - splits_SNP2   number of times this varaible is the split varaible 
   - adj_p_value_bonferrroni correction of int_Pvalue
   - adj_p_value_fdr          fdr correction of int_Pvalue
   - adj_p_value_Ch2          fdr correction of Ch2

so

- Bonferrroni on numeric SNP regression picks 21
- fdr on numeric SNP regression picks 88 (containing the 21 selected  by Bonferrroni )
- $\chi^2$ test on SNP factor variables picks  121 (containing 68 of the 88 selected by fdr)

```{R, echo=FALSE, eval=FALSE}

index <- which(interaction_snps_counts$Freq>=2)  #640

lm_results  <-  data.frame(interaction_snps_counts[index,],SNP1_effect=0,SNP1_Pvalue=1,SNP2_effect=0,SNP2_Pvalue=1,int_effect=0,int_Pvalue=1,Ch2=1)
dim(lm_results)
lm_results$correlations <- 0
selectedSNPs <- "pheno"
if(class(dt$pheno)=='factor'){
  for(l in 1:nrow(lm_results)){
    SNP1 <- strsplit(as.character(lm_results[l,1])," ")[[1]][1]
    SNP2 <- strsplit(as.character(lm_results[l,1])," ")[[1]][2]
    selectedSNPs <- union(selectedSNPs,SNP1)
    selectedSNPs <- union(selectedSNPs,SNP2)
    ind1  <-  which(colnames(dt)==SNP1)
    ind2  <-  which(colnames(dt)==SNP2)
    s  <-  summary(glm(dt$pheno ~ dt[,ind1]*dt[,ind2],family = binomial(link="logit")))$coeff
#l<-31 
#                        Estimate Std. Error    z value     Pr(>|z|)
#(Intercept)            0.4202587 0.06693247   6.278846 3.410951e-10
#dt[, ind1]            -0.2898685 0.06767282  -4.283382 1.840739e-05
#dt[, ind2]            -1.0043861 0.07960751 -12.616726 1.707751e-36
#dt[, ind1]:dt[, ind2]  0.2321571 0.06325044   3.670443 2.421308e-04
   lm_results$correlations[l] <- cor(dt[,ind1],dt[,ind2])
    #same as 
    #s2  <-  summary(glm(dt$pheno ~ dt[,ind1] + dt[,ind2] + dt[,ind1]:dt[,ind2],family = binomial(link="logit")))
    #s2$coeff

    #as factors
    d1<- factor(dt[,ind1])
    d2<- factor(dt[,ind2])

    s2  <-  glm(dt$pheno ~ d1 +d2 +d1:d2,family = binomial(link="logit"))
    nested.model <- glm(dt$pheno ~ d1+d2, family= binomial(link="logit"))
    #anova(nested.model, s2, test="LRT") #or
    aa<- lrtest(nested.model, s2)
#Likelihood ratio test
#
#Model 1: dt$pheno ~ d1 + d2
#Model 2: dt$pheno ~ d1 + d2 + d2:d2
#  #Df  LogLik Df Chisq Pr(>Chisq)
#1   5 -2985.5                    
#2   5 -2985.5  0     0          1
# 
#H0: You should use the nested model.
#Ha: You should use the complex model.
#Pr(>Chisq)=1  so use the simple model

    lm_results$splits_SNP1  <- t2[SNP1]
    lm_results$splits_SNP2  <- t2[SNP2]
    lm_results$Ch2[l] <- aa$"Pr(>Chisq)"[2]
    lm_results$SNP1_effect[l]  <-  s[2,1]
    lm_results$SNP1_Pvalue[l]  <-  s[2,4]
    lm_results$SNP2_effect[l]  <-  s[3,1]
    lm_results$SNP2_Pvalue[l]  <-  s[3,4]
    lm_results$int_effect[l]  <-  s[4,1]
    lm_results$int_Pvalue[l]  <-  s[4,4]

  }
}

##############
names(dt)[grep("X6_3",names(dt))]
SNP1 #[1] "X6_32912392" #just the last ones to come out of  the calculation above
SNP2 #[1] "X7_25326942"
dim(All_Interactions[grep(SNP2,All_Interactions$X1),]) #2 2
dim(All_Interactions[grep(SNP2,All_Interactions$X2),]) #35  2
#interaction_snps_counts[grep("X7_25326942", interaction_snps_counts$interaction_snps),]
aa<- interaction_snps_counts[grep("X6_32912392", interaction_snps_counts$interaction_snps),]
aa[grep("X7_25326942", aa$interaction_snps),]
#              interaction_snps Freq
#282088 X6_32912392 X7_25326942    2

grep("X6_32912392", lm_results[,1])
lm_results[ 640,]
#             interaction_snps Freq SNP1_effect  SNP1_Pvalue SNP2_effect SNP2_Pvalue int_effect int_Pvalue       Ch2 correlations splits_SNP1 splits_SNP2
#282086 X6_32912392 X7_25326942    2  -0.3853136 2.289531e-12 -0.02027243   0.7972829  0.1774552 0.04946914 0.1475998  0.004923941          18          14

ind1 <- grep("X6_32912392",names(dt))
ind2 <- grep("X7_25326942",names(dt))

s  <-  summary(glm(dt$pheno ~ dt[, ind1]*dt[, ind2],family = binomial(link="logit")))$coeff
#                                                                             Estimate Std. Error    z value     Pr(>|z|)
#(Intercept)                                                               -0.25152373 0.04720520 -5.3283057 9.913319e-08
#dt[, grep("X6_32912392", names(dt))]                                      -0.38531361 0.05492235 -7.0156068 2.289531e-12
#dt[, grep("X7_25326942", names(dt))]                                      -0.02027243 0.07892247 -0.2568651 7.972829e-01
#dt[, grep("X6_32912392", names(dt))]:dt[, grep("X7_25326942", names(dt))]  0.17745517 0.09032977  1.9645259 4.946914e-02

d1<- factor(dt[,ind1])
d2<- factor(dt[,ind2])
s2  <-  glm(dt$pheno ~ d1 +d2 +d1:d2,family = binomial(link="logit"))
nested.model <- glm(dt$pheno ~ d1+d2, family= binomial(link="logit"))
aa<- lrtest(nested.model, s2)
#Likelihood ratio test
#Model 1: dt$pheno ~ d1 + d2
#Model 2: dt$pheno ~ d1 + d2 + d1:d2
#  #Df  LogLik Df  Chisq Pr(>Chisq)
#1   5 -3174.1                     
#2   9 -3170.7  4 6.7867     0.1476
#all looks right

##########
lm_results <- lm_results[order(lm_results$int_Pvalue, decreasing = F),]
lm_results$adj_p_value_bonferrroni <- lm_results$int_Pvalue*nrow(lm_results) # a bonferrroni correction?
sum(lm_results$adj_p_value_bonferrroni < 0.05) #[1] 21

lm_results$adj_p_value_fdr <- p.adjust(lm_results$int_Pvalue, method = "fdr")
sum(lm_results$adj_p_value_fdr < 0.05) #[1] 88

lm_results$adj_p_value_Ch2 <- p.adjust(lm_results$Ch2 , method = "fdr")
sum(lm_results$adj_p_value_Ch2 < 0.05) #[1] 121
length(intersect( which(lm_results$adj_p_value_fdr < 0.05), which(lm_results$adj_p_value_Ch2 < 0.05)))# [1] 68
#save(lm_results, file=paste0(dirOut,"/lm_results_fixed.Rdata"))# fixed in the sense that the self-split pairs have been removed
#load(file=paste0(dirOut,"/lm_results_fixed.Rdata"))

names(lm_results)
# [1] "interaction_snps"        "Freq"                    "SNP1_effect"             "SNP1_Pvalue"             "SNP2_effect"             "SNP2_Pvalue"             "int_effect"             
# [8] "int_Pvalue"              "Ch2"                     "correlations"            "splits_SNP1"             "splits_SNP2"             "adj_p_value_bonferrroni" "adj_p_value_fdr"        
#[15] "adj_p_value_Ch2"        
lm_results[1:2,]
#              interaction_snps Freq SNP1_effect  SNP1_Pvalue SNP2_effect  SNP2_Pvalue int_effect   int_Pvalue          Ch2 correlations splits_SNP1 splits_SNP2 adj_p_value_bonferrroni
#280954 X6_32308125 X6_32535726   14  -1.0932552 5.783667e-39  -0.7248879 5.844399e-39  0.6083855 3.488304e-16 1.303556e-15  -0.14839455          18          14            2.232514e-13
#280951 X6_32308125 X6_32423632    3  -0.9145729 2.243391e-31  -0.4717926 3.387561e-16  0.5723224 5.161507e-14 4.721907e-12   0.07261602          18          14            3.303364e-11
#       adj_p_value_fdr adj_p_value_Ch2
#280954    2.232514e-13    8.342758e-13
#280951    1.651682e-11    1.007340e-09

#so
#
#- Bonferrroni on numeric SNP regression picks 21
#- fdr on numeric SNP regression picks 88
#- $\chi^2$ test on SNP factor variables picks  121

```

```{R, echo=FALSE, eval=FALSE}
# some plots of the  p-values for the different methods
plot(lm_results$int_Pvalue,pch=16)
points(lm_results$adj_p_value_fdr, col="red",pch=16)
points(lm_results$adj_p_value_bonferrroni , col="green",pch=16)
points(lm_results$adj_p_value_Ch2 , col="blue",pch=16)
abline(h=0.05)
legend(200,1,c("p-values, ordered", "fdr corrected", "Bonferroni corrected", "Chi2 p-values" ),pch=16,col=c("black", "red", "green","blue"))
#some concordance between the regression p-values and the Chi^2 p-values
#Bonferrroni finds 21, "fdr" finds 88 and \Chi^2 with fdr finds 121
length(intersect(which(lm_results$adj_p_value_Ch2  < 0.05), which(lm_results$adj_p_value_fdr  < 0.05))) # 68

lm_results.temp <- lm_results[order(lm_results$adj_p_value_Ch2, decreasing = F),]
plot(lm_results.temp$int_Pvalue, col="black")
points(lm_results.temp$adj_p_value_fdr, col="red")
points(lm_results.temp$adj_p_value_bonferrroni , col="green")
points(lm_results.temp$adj_p_value_Ch2 , col="blue")
abline(h=0.05)
#same thing

lm_results.temp[22,]
 #              interaction_snps Freq SNP1_effect SNP1_Pvalue SNP2_effect  SNP2_Pvalue int_effect int_Pvalue          Ch2 correlations splits_SNP1 splits_SNP2
#180240 X1_161191571 X6_32771829    2  -0.0257612   0.6986353  -0.8204181 3.832185e-29 0.00295658  0.9636916 4.178291e-05  0.001961207          18          14
#       adj_p_value_bonferrroni adj_p_value_fdr adj_p_value_Ch2
#180240                616.7626       0.9821061     0.001215503

ind1  <-  which(colnames(dt)=="X1_161191571")
ind2  <-  which(colnames(dt)=="X6_32771829")
s  <-  summary(glm(dt$pheno ~ dt[,ind1]*dt[,ind2],family = binomial(link="logit")))
#as factors
d1<- factor(dt[,ind1])
d2<- factor(dt[,ind2])
s2  <-  glm(dt$pheno ~ d1 +d2 +d1:d2,family = binomial(link="logit"))
nested.model <- glm(dt$pheno ~ d1+d2, family= binomial(link="logit"))
aa <- lrtest(nested.model, s2)
#Likelihood ratio test  aa$"Pr(>Chisq)" 4.178291e-05
# so the adj_p_value_fdr  is not significant but the adj_p_value_Ch2 is

options(width=195)
lm_results[1:21,]

# the top imp SNPs (selected by RFlocafdr) and how often they appear in the top interactions (glm, selected by FDR)
aa<-rep(NA, 74)
for ( i in 1:74){
    aa[i] <- length(grep(names(var_importance)[i] ,lm_results$interaction_snps[1:88]))
    }
# proving what? The SNPs selected by RFlocafdr are not selected as interactions

aa<-rep(NA, 74)
for ( i in 1:74){
    aa[i] <- length(grep(names(var_importance)[i] ,lm_results$interaction_snps[1:88]))
    }
# proving what? The SNPs selected by RFlocafdr are not selected as interactions

for ( i in 1:74){
    aa[i] <- length(grep(imp[i],lm_results$interaction_snps[1:88]))
    }
aa
#[1] 34 22 11  3  6  4  6  8  4  2  3  2  3  0  3  2  2  0  2  1  0  0  0  1  3
#[26]  2  0  1  0  0  1  0  1  1  0  0  0  1  0  0  0  0  1  0  0  0  0  0  1  0
#[51]  0  0  0  0  0  1  0  0  0  0  0  0  0  0  0  1  0  0  0  0  0  0  0  0

```

# data set up

```{R, echo=FALSE, eval=FALSE}

# Assuming your count data is in a vector 'x'
x <- interaction_snps_counts$Freq

dim(interaction_snps_counts)[1] #[1] 291142
sum(interaction_snps_counts$Freq) # 294120

94485*(94485 -1)/2 - 291142
x2<-  c(rep(0,4463369228),x)

#table(x2) long run time!
#    0         1         2      3      4       5      6      7      8      9     10     11     12     13     14     15     16     17     18     22     23     25     26     27     31     34     41  51     52     54     62     63     64     78     79     90    100    113    144    466 
#4463369228  290502    418     89     29     22     14     11      7      5      4      3      3      3      4     1      1      1      1      2      1      1      1      2      1      1      1   1      1      1      1      2      1      1      1      1      1      1      1      1

# First, prepare your data
counts <- as.numeric(names(table(x)))  # Unique count values (1, 2, 3, ...)
freqs <- as.numeric(table(x))          # Frequencies of each count
n_zeros <- 4463369228                   # Number of zeros

# Recreate the data based on your table x2
counts0 <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14,15,  16,  17,  18,  22,  23,  25,  26,  27,  31,  34,  41, 51,  52,  54,  62,  63,  64,  78,  79,  90, 100, 113, 144, 466) 
freqs0 <- c(4463369228, 290502, 418, 89, 29, 22, 14, 11, 7, 5, 4, 3, 3, 3, 4, 1, 1, 1, 1, 2, 1, 1, 1, 2, 1, 1, 1, 1,   1,   1,   1,   2,   1,   1,   1,   1,   1,   1,   1,   1)
# Recreate the data based on your table x
counts <- c( 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14,15,  16,  17,  18,  22,  23,  25,  26,  27,  31,  34,  41, 51,  52,  54,  62,  63,  64,  78,  79,  90, 100, 113, 144, 466) 
freqs <- c( 290502, 418, 89, 29, 22, 14, 11, 7, 5, 4, 3, 3, 3, 4, 1, 1, 1, 1, 2, 1, 1, 1, 2, 1, 1, 1, 1,   1,   1,   1,   2,   1,   1,   1,   1,   1,   1,   1,   1)

# Create a representative sample of this distribution (for computational efficiency)
set.seed(123)
sample_size <- 100000000
sample_probs <- freqs / sum(freqs)
x_sample <- sample(counts, size = sample_size, replace = TRUE, prob = sample_probs)

```

# Null distribution

What would the distribution of counts look like if 291142 interactions were assigned randomly?
Under a random assignment, this follows a binomial distribution that can be approximated by a Poisson distribution with:
$\lambda= 291,142 / 4,463,565,886 \approx 0.000065$

-  Count = 1: approximately 290,849 pairs
-  Count = 2: approximately 9.5 pairs
-  Count $\geq 3$: effectively 0 pairs

The observed data shows:

- Count = 1: 290502  pairs
-  Count = 2: 418   pairs
-  Count = 3: 89  pairs
-  Count $\geq 4$: 133 pairs

\clearpage
\FloatBarrier

# empiricalzip

- There are  4463613128 potential interactions. We have non-zero values for 294120, so there are a lot of 0 values.
- what to do to handle this? 
- **gpEMAlgorithm** fails -- set which estimates to calculate in the function call
- better name than **RealDataFDR**, what other data would we be running it on
- docuement output better
- "X6_32222493 X6_32771829" see below, how does it have  splits_SNP1=18,  splits_SNP2=14 but Freq=113

   - it doesnt t2["X6_32222493"] =202 t2["X6_32771829"]= 1532 -- lm_result has a problem

## some investigations of empiricalzip. 
```{R, echo=FALSE, eval=FALSE}
#devtools:::install_github("parsifal9/empiricalzip")
estimate<-  zigpEMAlgorithm(x2) #works
estimate<- zipEMAlgorithm(x2) #works
system.time(estimate<- pEMAlgorithm(x2)) ##works
estimate<- gpEMAlgorithm(x2) #Error in while (((abs(hold.lambda - lambdaEstimate)) > epsilon) || ((abs(hold.theta -  :   missing value where TRUE/FALSE needed

aa <- my.RealDataFDR(x2, 0.1) #has gpEMAlgorithm commented out, accepts data ratehr than a file
# Rejection counts
#       ZIGP ZIP   GP  P         ZIGP ZIP   GP  P        ZIGP  ZIP     GP  P
#   one stage procedure           two stage             Storey's FDR
#[1,]   50   82   82   82      57   82  104   82        35    53    53    53
#[2,]   50   82   82   82      57   82  104   82        35    53    53    53

#[[2]]Parameter estimates using cutoff method C_0

    ## eta: Zero-inflation parameter (proportion of structural zeros)
    ## lambda: Mean parameter
    ## theta: Dispersion parameter (NA for models that don't use it)
    ## pi: Estimated proportion of null hypotheses ($\pi_0$)
    ## C: The estimated cutoff value (\hat{C}) - positions with counts $\leq$ C are assumed to be from null
    ## AD: The truncation point (D_N) - positions with counts $\geq$ AD are automatically considered significant

#        eta         lambda    theta      pi          C0   AD
#[1,] 0.472686094 1.0020464 1.133941e-12 0.9999401    7    8   zigp
#[2,] 0.001277773 0.3350806           NA 0.9999075    5   14   zip
#[3,]          NA 0.3346515 0.000000e+00 0.9999072    4    5   gp
#[4,]          NA 0.3346515           NA 0.9999072    4   14    p
#
#[[3]]
#           [,1]      [,2]         [,3]      [,4] [,5] [,6]
#[1,] 0.47268609 1.0020464 1.133941e-12 0.9999401    7    8
#[2,] 0.00128352 0.3350822           NA 0.9999074    4   14
#[3,]         NA 0.3346515 0.000000e+00 0.9999072    4    5  
#[4,]         NA 0.3346515           NA 0.9999072    4   14

#empiricalzip  ZIGP one stage procedure
#empiricalzip pickes the interactions on the Freq, sointeraction_snps_counts[1:50,]
interaction_snps_counts$interaction_snps[1:50]

sum(freqs[-c(1:7)]) #[1] 68
sum(freqs[-c(1:8)]) #[1] 57
sum(freqs[-c(1:9)]) #[1] 50

```
The ZIGP model selected by epiricalzip has

- $\eta= 0.472686094$      Zero-inflation parameter (proportion of structural zeros)   
- $\lambda= 1.0020464$    Mean parameter
- $\theta= 1.133941e-12$    Dispersion parameter (NA for models that don't use it)  
- $\pi= 0.9999401$       Estimated proportion of null hypotheses

See table \@ref(tab:zigp_comparison)

\begin{table}
\caption{\label{tab:zigp_comparison} Observed vs. Expected Counts Under ZIGP Model selected by mpiricalzip}
\center
\begin{tabular}{lrr}
\hline
\textbf{Count} & \textbf{Observed} & \textbf{Expected (ZIGP)} \\
\hline
0 & 4,463,369,228 & 4,463,369,228 \\
1 & 290,502 & 290,490 \\
2 & 418 & 427 \\
3 & 89 & 85 \\
4 & 29 & 21 \\
5 & 22 & 5 \\
6-10 & 41 & 3 \\
11-20 & 13 & $<$1 \\
$>$20 & 22 & $\approx$0 \\
\hline
\textbf{Total} & 4,463,660,370 & 4,463,660,370 \\
\hline
\end{tabular}
\label{tab:zigp_comparison}
\end{table}

```{R, echo=FALSE, eval=FALSE}
# protein example (move this)

the protein exaple from @gauranEmpiricalNullEstimation2018
library(empiricalzip)
data(protein, package="empiricalzip")
k<-1
x <- protein[,(k+1)][!is.na(protein[,(k+1)])]
M <- length(x)
cf <- 10^(-10)
count <- c(length(x[which(x==0)]), tabulate(x, nbins = max(x)))
names(count) <- c(0:max(x))
aa <- my.RealDataFDR(x, 0.1) #
##     ZIGP ZIP   GP  P         ZIGP ZIP   GP  P        ZIGP  ZIP     GP  P
##      one stage procedure      two stage             Storey's FDR
## [1,]  191  212  212  212      201  212  223  212     200   211   211   211
## [2,]  143  205  212  212      143  205  229  212     162   204   211   211

## [[2]]
## #        eta      lambda    theta      pi    C0   AD   
## [1,] 0.3245463 1.9167862 0.1415783 0.4576160    6    7
## [2,] 0.2288849 1.0948534        NA 0.3985044    3    6
## [3,]        NA 0.8082199 0.0000000 0.3944174    3    4
## [4,]        NA 0.8082199        NA 0.3944174    3    6

## [[3]]
##           [,1]      [,2]      [,3]      [,4] [,5] [,6]
## [1,] 0.2251642 2.1452087 0.5741445 0.6141275    4   36
## [2,] 0.2759650 1.3856234        NA 0.4244176    2    6
## [3,]        NA 0.7994067 0.0000000 0.3929166    2    3
## [4,]        NA 0.7994067        NA 0.3929166    2    6

# ZIGP parameters
eta <- 0.3245463
lambda <- 1.9167862 
theta <- 0.1415783

# Generate the plots and analysis
results <- plot_zigp_comparison(
  observed_counts = count,
  eta = eta,
  lambda = lambda, 
  theta = theta,
  max_display = 30,
  y_min = -15  # Extended y-axis
)

results$log_scale_tail
abline(v=6)

plot_normalized_mixture(
   eta = eta,
   lambda =  lambda, 
   theta = theta,
   pi0 = 1,
   n_binom = 15,
   p_binom = 0.2,
   shift = 7
)

```

```{R, echo=FALSE, eval=FALSE}

# Create a named vector with the observed counts
observed_data <- freqs
names(observed_data) <- as.character(counts)

# ZIGP parameters
eta <- 0.9993963  
lambda <- 1.0020464
theta <- 1.133941e-12

# Generate the plots and analysis
results <- plot_zigp_comparison(
  observed_counts = observed_data,
  eta = eta,
  lambda = lambda, 
  theta = theta,
  max_display = 30,
  y_min = -15  # Extended y-axis
)

# Display the plots
library(gridExtra)
grid.arrange(results$log_scale, results$linear_scale, ncol = 1)

# Testing separation between null and alternative distributions
# First, examine where the residuals cross from positive to negative
crossover_point <- min(which(results$separation_analysis$Residual < 0))
print(paste("Crossover point where observed becomes less than expected:", 
            results$separation_analysis$Count[crossover_point]))

# Calculate patterns of overdispersion
counts_with_excess <- sum(results$separation_analysis$Observed > results$separation_analysis$Fitted)
print(paste("Number of count values with excess observations:", counts_with_excess))

# Calculate the count where maximum deviation occurs
max_positive_deviation <- which.max(results$separation_analysis$RelativeDiff)
print(paste("Count with maximum positive deviation:", 
            results$separation_analysis$Count[max_positive_deviation]))

# Save the improved plot
#ggsave("./script3/zigp_log_scale_extended.png", results$log_scale, width = 10, height = 6)
#ggsave("zigp_ratio_plot.png", results$ratio_plot, width = 10, height = 6)

```

```{r zigplogscaleextended, eval=TRUE, echo=FALSE, fig.cap='empiricalzip fitted ZIPG',  fig.align = 'center', out.width = '50%'}
knitr::include_graphics("./script3/zigp_log_scale_extended.png")

```

\clearpage
\FloatBarrier
## Local FDR Calculation
The local FDR (false discovery rate) at a particular count value is calculated as:

- fdr(t) = ($\pi_0 \times f_0(t)) / f(t)$

Where:

-  $t4 is the mutation count
-  $\pi_0$ is the estimated proportion of nulls
-  $ f_0(t)$ is the null distribution probability at count $t$
-  $f(t)$ is the observed mixture distribution probability at count $t$

In the my.RealDataFDR function, this calculation happens here:

```{verbatim}
LFDR <- pi[model, CO[method, model]] * gpdMixture(x = sort(x), 
    estimate[model, 1, CO[method, model]], estimate[model, 
    2, CO[method, model]], estimate[model, 3, CO[method, 
    model]]) * f12(sort(x), length(x))
```
Breaking this down:

-  pi[model, CO[method, model]] is the estimated $\pi_0$  (proportion of nulls)
-  gpdMixture() calculates  $ f_0(t)$ , the probability under the null distribution (ZIGP, ZIP, GP, or P) using the estimated parameters
-  f12(sort(x), length(x)) is a correction factor related to f(t). Looking at the f12 function in the RcppExports.R, this factor is calculated as:

```{verbatim}
   M / length(x[which(x==x[i])])
```
where M is the total number of positions. This is effectively equivalent to $1/f(t)$, where $f(t)$ is the empirical probability of observing count $t$.

For the one-stage procedure, positions are declared significant if:

```{verbatim}
sum(LFDR[!is.na(LFDR)] < FDRLevel)
```
meaning their local FDR is below the specified threshold (in your case, 0.1).

Why Some Positions with Counts >7 Aren't Called Significant

With the ZIGP model estimated parameters for your data, the local FDR for positions with counts of 8, 9, etc. depends on:

-   How rare such counts are in the data (affecting $f(t)$)
-  How probable they are under the null distribution (affecting $f_0(t)$)

When a position has a count that is high but not extremely high relative to the estimated null distribution, its local
FDR might still exceed the threshold, especially when $\pi_0$ is estimated to be very close to 1.0.

This is why 7 positions with counts >7 aren't called significant by the one-stage procedure - their local FDR values
exceed your threshold of 0.1, indicating there's still a substantial probability they come from the null distribution
despite their relatively high counts.

The two-stage procedure, by contrast, automatically declares positions with counts $\geq$ AD (in this case 8) as significant,
bypassing the local FDR calculation for these high counts.

1. Stage 1 (Screening Step): We reject the $H_0 j \textrm{if } j \leq D_N(\hat{C})$ and do not reject if $j \leq \hat{C}$.
2. Stage 2 (Testing step): For $\hat{C} + 1 \leq j < D_N(\hat{C})$ , we calculate the local FDR

## Simplified two-stage procedure for empiricalzip

-  "Estimated C: 1"
-  "D_N threshold: 3"
-  "Significant counts:"   3  4  5  7 27
- used x_sample so some counts missing
- could use this value

```{R, echo=FALSE, eval=FALSE}
# Simplified two-stage procedure for extreme zero-inflation
simplified_two_stage <- function(data, fixed_C = 1, alpha = 0.05) {
  # Extract the counts and their frequencies
  tab <- table(data)
  counts <- as.numeric(names(tab))
  freqs <- as.numeric(tab)
  N <- sum(freqs)

  # Given extreme zero-inflation, set C directly
  C_hat <- fixed_C

  # Estimate parameters for the null distribution (ZIGP)
  # Given the extreme zeros, simplify parameter estimation
  eta_est <- freqs[counts == 0] / N  # Zero-inflation parameter

  # For non-zero values, estimate lambda and theta
  non_zero_counts <- rep(counts[counts > 0], freqs[counts > 0])
  lambda_est <- mean(non_zero_counts)
  theta_est <- min(0.5, var(non_zero_counts) / lambda_est - 1)

  # Calculate D_N (direct approach for computational stability)
  # With extreme zeros, a lower D_N is appropriate
  D_N <- max(3, C_hat + 1)  # Simple heuristic

  # Create results structure
  results <- list(
    C_hat = C_hat,
    D_N = D_N,
    parameters = c(eta = eta_est, lambda = lambda_est, theta = theta_est),
    significant_counts = counts[counts >= D_N]
  )

  return(results)
}

# Load necessary libraries
library(VGAM)
library(stats)

# Function to implement the two-stage procedure from Gauran et al.
two_stage_zigp_procedure <- function(data, alpha = 0.05) {
  # Extract the counts and their frequencies
  counts <- as.numeric(names(table(data)))
  freqs <- as.numeric(table(data))
  N <- sum(freqs)

  # Calculate the empirical probabilities
  f_hat <- freqs / N

  # Define function to estimate C using the likelihood approach (C1 method)
  estimate_C1 <- function(counts, freqs, N) {
    # Initialize variables
    max_S <- -Inf
    best_C <- 1

    # Try different values of C
    for (v in 1:(max(counts)-1)) {
      # Subset data for estimation
      subset_counts <- counts[counts <= v]
      subset_freqs <- freqs[counts <= v ]

      # Fit ZIGP to data up to C using MLE or method of moments
      # This is a simplified approach - in practice, use EM algorithm
      p_zeros <- subset_freqs[subset_counts == 0] / sum(subset_freqs)
      lambda_est <- mean(rep(subset_counts, subset_freqs))

      # Due to extreme zero-inflation, we'll use high eta
      eta_est <- max(0.95, p_zeros - (1-p_zeros)*exp(-lambda_est))
      theta_est <- min(0.5, var(rep(subset_counts, subset_freqs)) / lambda_est - 1)

      # Calculate pi_0 estimate
      pi0_est <- min(1, sum(freqs[counts <= v]) / (N * sum(zigp_pmf(0:v, eta_est, lambda_est, theta_est))))

      # Calculate S_v statistic
      S_v <- sum(freqs[counts <= v] * log(zigp_pmf(counts[counts <= v], eta_est, lambda_est, theta_est) / f_hat[counts <= v]))

      # Update if this is the best C so far
      if (S_v > max_S) {
        max_S <- S_v
        best_C <- v
        best_params <- c(eta_est, lambda_est, theta_est, pi0_est)
      }
    }

    return(list(C = best_C, params = best_params))
  }

  # Define ZIGP PMF function (simplified)
  zigp_pmf <- function(x, eta, lambda, theta) {
    result <- numeric(length(x))

    # Handle zeros separately due to zero-inflation
    result[x == 0] <- eta + (1 - eta) * exp(-lambda)

    # Handle positive values
    for (i in which(x > 0)) {
      val <- x[i]
      prob <- (1 - eta) * (lambda * (lambda + theta * val)^(val - 1) * exp(-lambda - theta * val)) / factorial(val)
      result[i] <- prob
    }

    return(result)
  }

  # Calculate D_N (truncation point)
  calculate_D_N <- function(C, params) {
    eta <- params[1]
    lambda <- params[2]
    theta <- params[3]

    # Calculate D_N based on the paper's formula
    # Simplified version for extreme zero-inflation
    D1 <- ceiling(log(N) / (-log(theta)))
    D2 <- ceiling(C + 1)

    return(max(D1, D2))
  }

  # Stage 1: Estimate C and calculate D_N
  C_result <- estimate_C1(counts, freqs, N)
  C_hat <- C_result$C
  params <- C_result$params
  D_N <- calculate_D_N(C_hat, params)

  # Stage 2: Apply local FDR to values between C+1 and D_N-1
  local_fdr_result <- numeric(max(counts))

  for (j in (C_hat+1):(D_N-1)) {
    if (j %in% counts) {
      # Calculate local FDR
      null_density <- zigp_pmf(j, params[1], params[2], params[3])
      mixture_density <- f_hat[counts == j]
      local_fdr_result[j] <- params[4] * null_density / mixture_density
    }
  }

  # Determine significant counts
  significant_counts <- numeric(0)

  # Reject all counts \geq D_N
  if (any(counts >= D_N)) {
    significant_counts <- c(significant_counts, counts[counts >= D_N])
  }

  # Reject counts between C+1 and D_N-1 if local FDR \leq alpha
  for (j in (C_hat+1):(D_N-1)) {
    if (j %in% counts && local_fdr_result[j] <= alpha) {
      significant_counts <- c(significant_counts, j)
    }
  }

  return(list(
    C_hat = C_hat,
    D_N = D_N,
    parameters = params,
    local_fdr = local_fdr_result,
    significant_counts = significant_counts
  ))
} #end two_stage_zigp_procedure

## Apply the simplified two-stage procedure
results <- simplified_two_stage(x_sample, fixed_C = 1)

# Display results
print(paste("Estimated C:", results$C_hat))
print(paste("D_N threshold:", results$D_N))
print("Significant counts:")
print(results$significant_counts)

```

# hurdle model implementation

```{R, echo=FALSE, eval=FALSE}

library(pscl)  # For hurdle model implementation

# Fit a hurdle model
hurdle_model <- hurdle(
  x ~ 1,               # Formula (just intercept in this simple case)
  data = data.frame(x = x_sample),
  dist = "negbin",     # Distribution for non-zero counts
  zero.dist = "binomial" # Distribution for zero vs non-zero process
)

summary(hurdle_model)

```

```{R, echo=FALSE, eval=FALSE}
# what is this? remove it
# Custom function for extreme zero-inflation
extreme_zero_model <- function(x, min_count = 3) {
  # Calculate basic statistics
  zeros <- sum(x == 0)
  n <- length(x)
  p_zero <- zeros/n

  # Estimate significance threshold
  # For extreme zero-inflation, we can use empirical approach
  unique_counts <- sort(unique(x))
  freq_table <- table(x)

  # Simple rule: counts are significant when they occur with 
  # frequency less than 0.1% of non-zero observations
  rarity_threshold <- 0.001 * (n - zeros)
  significant <- unique_counts[unique_counts >= min_count & 
                              freq_table[as.character(unique_counts)] < rarity_threshold]

  return(list(
    p_zero = p_zero,
    significant_counts = significant
  ))
}

```

# what is this 
```{R, echo=FALSE,eval=FALSE}

length(intersect(interaction_snps_counts$interaction_snps[1:50],lm_results$interaction_snps[1:88])) #22
# lm_results (has been ordered by lm_results$int_Pvalue)

t1 <- setdiff(interaction_snps_counts$interaction_snps[1:50],lm_results$interaction_snps[1:88])
match(t1,interaction_snps_counts$interaction_snps)  #3  4  5  7

t1[1]
#1] "X6_32222493 X6_32771829"
interaction_snps_counts[3,]
#              interaction_snps Freq
#280884 X6_32222493 X6_32771829  113

lm_results[grep("X6_32771829", lm_results$interaction_snps),] #301
lm_results[grep("X6_32222493 X6_32771829", lm_results$interaction_snps),] 
#              interaction_snps Freq SNP1_effect  SNP1_Pvalue SNP2_effect  SNP2_Pvalue int_effect int_Pvalue        Ch2 correlations splits_SNP1 splits_SNP2
#280884 X6_32222493 X6_32771829  113   0.3937617 6.972158e-09  -0.6516203 2.576182e-27 -0.1303979 0.08982739 0.03193039   -0.3742329          18          14
#       adj_p_value_bonferrroni adj_p_value_fdr adj_p_value_Ch2
#280884                57.48953       0.2521143       0.1011656

# so this is not significant for any of the tests except empiricalzip
# how does it have  splits_SNP1=18,  splits_SNP2=14 but Freq=113

chisq.test(table( dt[,"X6_32222493",], dt[,"X6_32771829",]))

aa<-rep(NA, 74)
for ( i in 1:74){
    aa[i] <- length(grep(imp[i],interaction_snps_counts$interaction_snps[1:50]))
    }
aa
# [1] 21 13  8  7  5  5  5  8  5  3  2  2  2  1  2  2  1  1  0  0  1  0  0  0  0  0  1  0  0  0  1  0  0  1  1  0  0  0  0  0  0  0  0  0  0  0  0  0  0  0  0  0  1  0  0
#[56]  0  0  0  0  0  0  0  0  0  0  0  0  0  0  0  0  0  0  0

index <- match(selectedSNPs,colnames(dt))
index <- index[order(index)]

data4interactions <- dt[,index]
saveRDS(data4interactions,paste0(dirOut,"/data4interactions.rds"))

SNPs <- lm_results$interaction_snps

SNPs <- as.character(SNPs)
SNPs <- gsub("X","" ,SNPs)
#SNPs <- read.table("Results/SNPs.txt",header=T)

```

4. what is the intesection of imp and interacting transcripts

   - see table at end of section

- "X6_32222493 X6_32771829" see below, how does it have  splits_SNP1=18,  splits_SNP2=14 but Freq=113

   - it doesnt t2["X6_32222493"] =202 t2["X6_32771829"]= 1532 -- lm_result has a problem

# compare some distributions

- modified compare_distributions function and related plotting functions to work with a frequency table input rather than requiring the full vector. 
This will be much more memory-efficient when dealing with large datasets that have many zeros.

```{R, echo=FALSE, eval=FALSE}
# Then run the comparison
results <- compare_distributions_from_freq(counts, freqs, n_zeros)

# View results
print(results$comparison)
#    Model LogLikelihood Parameters     AIC     BIC AIC_Ratio
#2 ZI-Zipf      -3102980          2 6205964 6206005         1
#3    ZINB          -Inf          3     Inf     Inf       Inf
#1    ZIGP           NaN          3     NaN     NaN       NaN

print(paste("Best model:", results$best_model))

#png("./script3/ZI_Zipf_plot.png")
results$plots$"ZI_Zipf"
#dev.off()
#png("./script3/ZIGP_plot.png")
results$plots$"ZIGP"
#dev.off()
#png("./script3/ZINB_plot.png")
results$plots$"ZINB"
#dev.off()

results$parameters
#$ZIGP $ZIGP$eta [1] 0.9999 $ZIGP$lambda [1]  0.9752822  $ZIGP$theta  [1]  0.03459262
#$ZI_Zipf $ZI_Zipf$p0 [1] 0.9999348 $ZI_Zipf$alpha [1] 8.288837 
#$ZINB $ZINB$p0 [1] 0.9999348 $ZINB$pi_0  [1] 0.9998953 $ZINB$size  [1] 13.84857 $ZINB$mu [1] 1.010229

```

```{R, zipfplot, echo=FALSE, eval=TRUE, fig.cap="Zero Inflated Power Law (Zipf) Distribution", out.width = '50%',fig.align = 'center'}
knitr::include_graphics("./script3/ZI_Zipf_plot.png")
```

```{R, zigpplot, echo=FALSE, eval=TRUE, fig.cap="Zero Inflated Generalized Poisson", out.width = '50%',fig.align = 'center'}
knitr::include_graphics("./script3/ZIGP_plot.png")
```

```{R, zinbplot, echo=FALSE, eval=TRUE, fig.cap="Zero Inflated Negative Binomial", out.width = '50%',fig.align = 'center'}
knitr::include_graphics("./script3/ZINB_plot.png")
```

```{R, echo=FALSE, eval=FALSE}
# Modified functions to work with frequency data instead of raw vectors

# ZIGP probability mass function
zigp_pmf <- function(x, eta, lambda, theta) {
  result <- numeric(length(x))

  # Handle zeros
  result[x == 0] <- eta + (1 - eta) * exp(-lambda)

  # Handle positive values
  for (i in which(x > 0)) {
    val <- x[i]
    prob <- (1 - eta) * lambda * (lambda + theta * val)^(val - 1) * exp(-lambda - theta * val) / factorial(val)
    result[i] <- prob
  }

  return(result)
}

# Modifications to handle extreme zero-inflation
# For the ZIGP model:
estimate_zigp_params_from_freq <- function(counts, freqs, n_zeros) {
  # Calculate basic statistics
  n <- sum(freqs) + n_zeros
  p_zeros <- n_zeros / n

  # If extreme zero-inflation
  if (p_zeros > 0.9999) {
    # Use a simpler model with high but not exactly 1 eta
    return(list(eta = 0.9999, lambda = mean(counts[counts > 0]), theta = 0))
  }

    # Rest of the function as before...
  # Calculate statistics for positive values
  if (sum(freqs[counts > 0]) == 0) {
    return(list(eta = 1, lambda = 1, theta = 0))  # All zeros case
  }

  # Calculate mean and variance of positive values
  # Weighted calculations using frequencies
  mean_pos <- sum(counts[counts > 0] * freqs[counts > 0]) / sum(freqs[counts > 0])

  # Calculate variance
  var_pos <- sum(freqs[counts > 0] * (counts[counts > 0] - mean_pos)^2) / sum(freqs[counts > 0])

  # Estimate parameters using method of moments
  # For zero-inflated data, first estimate eta
  eta_est <- max(0, (p_zeros - exp(-mean_pos)) / (1 - exp(-mean_pos)))

  # Then estimate lambda and theta from positive values
  lambda_est <- mean_pos
 # Estimate theta from variance relationship
  # In GP: Var(X) = lambda / (1-theta)^2
  # Solving for theta
  theta_est <- max(0, min(0.999, 1 - sqrt(lambda_est / var_pos)))

  return(list(eta = eta_est, lambda = lambda_est, theta = theta_est))
}
aa<- estimate_zigp_params_from_freq(counts, freqs, n_zeros)

```

```{R, echo=FALSE, eval=FALSE}
# Improved function to estimate ZIGP parameters from frequency data
# This handles zero-inflation properly by treating zeros separately
improved_estimate_zigp_params <- function(counts, freqs, n_zeros) {
  # Calculate basic statistics
  n <- sum(freqs) + n_zeros
  p_zeros <- n_zeros / n

  # Focus only on non-zero counts for fitting Generalized Poisson
  pos_idx <- counts > 0
  pos_counts <- counts[pos_idx]
  pos_freqs <- freqs[pos_idx]

  # Calculate weighted statistics for positive counts
  mean_pos <- sum(pos_counts * pos_freqs) / sum(pos_freqs)
  var_pos <- sum(pos_freqs * (pos_counts - mean_pos)^2) / sum(pos_freqs)

  # Check if the data is actually highly overdispersed
  if (var_pos < mean_pos) {
    # If variance is less than mean (underdispersed), default to Poisson (theta = 0)
    lambda_est <- mean_pos
    theta_est <- 0
  } else {
    # For overdispersed data, estimate theta using variance relationship
    # In GP: Var(X) = lambda * (1 + theta)^2 / (1 - theta)^2
    # Solving for theta (simplified version)
    dispersion_ratio <- var_pos / mean_pos

    if (dispersion_ratio <= 1) {
      # No overdispersion
      theta_est <- 0
    } else if (dispersion_ratio < 10) {
      # Moderate overdispersion - using approximation
      theta_est <- 1 - 1/sqrt(dispersion_ratio)
      theta_est <- max(0, min(0.8, theta_est))  # Cap at 0.8 for stability
    } else {
      # Extreme overdispersion - use caution
      theta_est <- 0.8  # Cap at 0.8 for numerical stability
    }

    # Adjust lambda based on theta estimate
    lambda_est <- mean_pos * (1 - theta_est)
  }

  # Calculate eta (zero-inflation parameter) using the fitted GP distribution
  p_zero_gp <- exp(-lambda_est)  # Probability of zero under GP

  # Handle extreme zero-inflation
  if (p_zeros > 0.9999) {
    eta_est <- 0.9999  # Cap at 0.9999 for numerical stability
  } else if (p_zeros <= p_zero_gp) {
    # If observed zeros are fewer than expected from GP alone
    eta_est <- 0  # No zero-inflation needed
  } else {
    # Calculate eta using the zero-inflation formula
    eta_est <- (p_zeros - p_zero_gp) / (1 - p_zero_gp)
    eta_est <- max(0, min(0.9999, eta_est))  # Ensure it's in [0, 0.9999]
  }

  # Print debugging information
  cat("Estimated parameters:\n")
  cat("Mean of positive counts:", mean_pos, "\n")
  cat("Variance of positive counts:", var_pos, "\n")
  cat("Dispersion ratio (var/mean):", var_pos/mean_pos, "\n")
  cat("eta:", eta_est, "\n")
  cat("lambda:", lambda_est, "\n")
  cat("theta:", theta_est, "\n")

  # Return estimates
  return(list(eta = eta_est, lambda = lambda_est, theta = theta_est))
}

# Function to calculate ZIGP probabilities
zigp_pmf <- function(x, eta, lambda, theta) {
  result <- numeric(length(x))

  # Handle zeros separately due to zero-inflation
  result[x == 0] <- eta + (1 - eta) * exp(-lambda)

  # Handle positive values
  for (i in which(x > 0)) {
    val <- x[i]
    prob <- (1 - eta) * lambda * (lambda + theta * val)^(val - 1) * 
            exp(-lambda - theta * val) / factorial(val)
    result[i] <- prob
  }

  return(result)
}

# Function to compare observed vs expected frequencies
compare_observed_expected <- function(counts, freqs, n_zeros, params) {
  # Create full data frame for comparison
  all_counts <- c(0, counts)
  all_freqs <- c(n_zeros, freqs)
  total_n <- sum(all_freqs)

  # Calculate ZIGP probabilities
  expected_probs <- zigp_pmf(all_counts, params$eta, params$lambda, params$theta)
  expected_freqs <- expected_probs * total_n

  comparison <- data.frame(
    Count = all_counts,
    Observed = all_freqs,
    Expected = expected_freqs,
    ObsProb = all_freqs / total_n,
    ExpProb = expected_probs,
    Ratio = all_freqs / expected_freqs
  )

  return(comparison)
}

# Example usage:
# counts <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
# freqs <- c(290502, 418, 89, 29, 22, 14, 11, 7, 5, 4)
# n_zeros <- 4463369228
# 
# # Estimate improved parameters
params <- improved_estimate_zigp_params(counts, freqs, n_zeros)
# 
# # Compare observed vs expected
comparison <- compare_observed_expected(counts, freqs, n_zeros, params)
print(comparison)
estimate_zigp_params_from_freq<- improved_estimate_zigp_params

```

```{R, echo=FALSE, eval=FALSE}
# Power law PMF (Zipf distribution)
zipf_pmf <- function(x, alpha, xmin = 1) {
  result <- numeric(length(x))
  # Use a more efficient approach for normalization
  denom <- zeta(alpha, xmin)  # Riemann zeta function starting at xmin

  result[x < xmin] <- 0
  result[x >= xmin] <- (x[x >= xmin]^(-alpha)) / denom

  return(result)
}

# Helper function: approximate Riemann zeta function
zeta <- function(alpha, xmin = 1) {
  # For alpha > 1, we can use a finite sum for approximation
  max_k <- 10000  # This should be enough for good approximation with alpha > 2
  sum((xmin:max_k)^(-alpha))
}

# For zero-inflated data, combine with point mass at zero
zi_zipf_pmf <- function(x, p0, alpha, xmin = 1) {
  result <- numeric(length(x))
  result[x == 0] <- p0
  result[x > 0] <- (1 - p0) * zipf_pmf(x[x > 0], alpha, xmin)
  return(result)
}

# Zero-inflated negative binomial
zinb_pmf <- function(x, p0, size, mu) {
  result <- numeric(length(x))
  result[x == 0] <- p0 + (1 - p0) * dnbinom(0, size = size, mu = mu)
  result[x > 0] <- (1 - p0) * dnbinom(x[x > 0], size = size, mu = mu)
  return(result)
}

# Fit Zipf distribution from frequency data
fit_zipf_from_freq <- function(counts, freqs) {
  # Skip 0 counts
  pos_idx <- counts > 0
  if(sum(pos_idx) == 0) return(list(alpha = 1))

  pos_counts <- counts[pos_idx]
  pos_freqs <- freqs[pos_idx]

  # MLE for alpha (approximate)
  alpha_start <- 1.5  # Starting value

  neg_ll <- function(alpha) {
    # Calculate log likelihood weighted by frequencies
    -sum(pos_freqs * log(zipf_pmf(pos_counts, alpha)))
  }

  result <- optim(alpha_start, neg_ll, method = "BFGS")
  return(list(alpha = result$par))
}

# Fit Negative Binomial from frequency data
fit_nb_from_freq <- function(counts, freqs) {
  # Skip 0 counts
  pos_idx <- counts > 0
  if(sum(pos_idx) == 0) return(list(size = 1, mu = 1))

  pos_counts <- counts[pos_idx]
  pos_freqs <- freqs[pos_idx]

  # Calculate mean and variance using frequencies as weights
  mu_est <- sum(pos_counts * pos_freqs) / sum(pos_freqs)

  # Calculate variance
  var_est <- sum(pos_freqs * (pos_counts - mu_est)^2) / sum(pos_freqs)

  if(var_est <= mu_est) var_est <- mu_est * 1.1  # Ensure overdispersion

  size_est <- mu_est^2 / (var_est - mu_est)

  return(list(size = size_est, mu = mu_est))
}

# Improved plotting function for frequency data
plot_distribution_from_freq <- function(counts, freqs, n_zeros, model_name, model_probs, p0 = NULL) {
  # Add the zero count
  all_counts <- c(0, counts)
  all_freqs <- c(n_zeros, freqs)

  # Calculate observed probabilities
  total_n <- sum(all_freqs)
  obs_probs <- all_freqs / total_n

  # Focus on counts with significant probability
  max_count <- min(50, max(all_counts))

  # Create a sequence of counts to display
  display_counts <- 0:max_count

  # Create data for plotting
  plot_data <- data.frame(
    Count = factor(display_counts),
    Observed = NA,
    Fitted = NA
  )

  # Fill in observed probabilities
  for(i in 1:length(all_counts)) {
    if(all_counts[i] <= max_count) {
      idx <- which(display_counts == all_counts[i])
      if(length(idx) > 0) {
        plot_data$Observed[idx] <- obs_probs[i]
      }
    }
  }

  # Fill in fitted probabilities
  model_counts <- c(0, counts) # Include zero
  if(length(model_probs) < length(model_counts)) {
    # Handle case where model_probs might not have same length
    model_probs <- c(model_probs[1], model_probs)  # Duplicate first value for zero
  }

  for(i in 1:length(model_counts)) {
    if(model_counts[i] <= max_count) {
      idx <- which(display_counts == model_counts[i])
      if(length(idx) > 0) {
        plot_data$Fitted[idx] <- model_probs[i]
      }
    }
  }

  # Convert to long format for plotting
  library(tidyr)
  plot_long <- pivot_longer(plot_data, 
                           cols = c("Observed", "Fitted"),
                           names_to = "Type", 
                           values_to = "Probability")

  # Remove missing values
  plot_long <- plot_long[!is.na(plot_long$Probability),]

  # Create the plot
  library(ggplot2)
  model_label <- paste("Fitted", model_name)
  fill_values <- c("red", "blue")
  names(fill_values) <- c("Observed", "Fitted")

  p <- ggplot(plot_long, aes(x = Count, y = Probability, fill = Type)) +
    geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
    scale_y_log10(limits = c(min(plot_long$Probability[plot_long$Probability > 0]) * 0.5, 1)) +
    theme_minimal() +
    labs(title = paste("Observed vs Fitted", model_name, "Probabilities (log scale)"),
         subtitle = if(!is.null(p0)) paste0("p0 = ", round(p0, 4)) else "",
         x = "Count Value", y = "Probability (log scale)") +
    scale_fill_manual(values = fill_values) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  return(p)
}

# Main function to compare distributions based on frequency data
compare_distributions_from_freq <- function(counts, freqs, n_zeros) {
  # Calculate total number of observations
    n <- sum(freqs) + n_zeros
    # For the log-likelihood calculations:
    # Add small epsilon to avoid log(0)
    epsilon <- 1e-10

  # Calculate empirical probability of zero
  p0_emp <- n_zeros / n

  # Fit ZIGP (as baseline)
  params_zigp <- estimate_zigp_params_from_freq(counts, freqs, n_zeros)

  # Calculate ZIGP probabilities and log-likelihood
  unique_counts <- sort(unique(c(0, counts)))
  probs_zigp <- zigp_pmf(unique_counts, params_zigp$eta, params_zigp$lambda, params_zigp$theta)

  # Calculate log likelihood using frequencies
  count_map <- match(c(0, counts), unique_counts)
  #ll_zigp <- n_zeros * log(probs_zigp[1]) + sum(freqs * log(probs_zigp[count_map[-1]]))
  ll_zigp <- n_zeros * log(max(probs_zigp[1], epsilon)) + sum(freqs * log(pmax(probs_zigp[count_map[-1]], epsilon)))

  # Fit power law (Zipf)
  params_zipf <- fit_zipf_from_freq(counts, freqs)
  p0_est <- p0_emp

  # Calculate Zipf probabilities and log-likelihood
  probs_zipf_pos <- zipf_pmf(counts, params_zipf$alpha)
  probs_zipf <- c(p0_est, (1-p0_est) * probs_zipf_pos)
  #ll_zipf <- n_zeros * log(p0_est) + sum(freqs * log((1-p0_est) * probs_zipf_pos))
  ll_zipf <- n_zeros * log(max(p0_est, epsilon)) +   sum(freqs * log(pmax((1-p0_est) * probs_zipf_pos, epsilon)))

  # Fit Negative Binomial
  params_nb <- fit_nb_from_freq(counts, freqs)

    # Calculate NB probabilities and log-likelihood
    # For the ZINB model:
    # Ensure the parameters are reasonable for extreme zero-inflation
    params_nb$size <- pmin(params_nb$size, 1000)  # Cap the size parameter

  probs_nb_pos <- dnbinom(counts, size = params_nb$size, mu = params_nb$mu)
  probs_nb_0 <- dnbinom(0, size = params_nb$size, mu = params_nb$mu)

  # For zero-inflated model
  pi_0 <- (p0_est - probs_nb_0) / (1 - probs_nb_0)
  pi_0 <- max(0, min(1, pi_0))  # Ensure valid probability

  probs_zinb <- c(p0_est, (1-pi_0) * probs_nb_pos)
  ll_nb <- n_zeros * log(p0_est) + sum(freqs * log((1-pi_0) * probs_nb_pos))

  # Calculate AIC and BIC
  aic_zigp <- -2 * ll_zigp + 2 * 3  # 3 parameters
  aic_zipf <- -2 * ll_zipf + 2 * 2  # 2 parameters
  aic_nb <- -2 * ll_nb + 2 * 3      # 3 parameters

  bic_zigp <- -2 * ll_zigp + log(n) * 3
  bic_zipf <- -2 * ll_zipf + log(n) * 2
  bic_nb <- -2 * ll_nb + log(n) * 3

  # Create comparison table
  results <- data.frame(
    Model = c("ZIGP", "ZI-Zipf", "ZINB"),
    LogLikelihood = c(ll_zigp, ll_zipf, ll_nb),
    Parameters = c(3, 2, 3),
    AIC = c(aic_zigp, aic_zipf, aic_nb),
    BIC = c(bic_zigp, bic_zipf, bic_nb),
    AIC_Ratio = c(aic_zigp/aic_zipf, 1, aic_nb/aic_zipf)  # Add ratio to better interpret differences
  )

  # Sort by AIC
  results <- results[order(results$AIC),]

  # Create plots for each model
  p_zigp <- plot_distribution_from_freq(counts, freqs, n_zeros, "ZIGP", probs_zigp, params_zigp$eta)
  p_zipf <- plot_distribution_from_freq(counts, freqs, n_zeros, "ZI-Zipf", probs_zipf, p0_est)
  p_nb <- plot_distribution_from_freq(counts, freqs, n_zeros, "ZINB", probs_zinb, p0_est)

  # Return results
  return(list(
    comparison = results,
    best_model = results$Model[1],
    plots = list(ZIGP = p_zigp, ZI_Zipf = p_zipf, ZINB = p_nb),
    parameters = list(
      ZIGP = params_zigp,
      ZI_Zipf = c(list(p0 = p0_est), params_zipf),
      ZINB = c(list(p0 = p0_est, pi_0 = pi_0), params_nb)
    )
  ))
}

```

\begin{table}[h]
\centering
\caption{Model Comparison for Rheumatoid Arthritis Interaction Data}
\begin{tabular}{lrrrrr}
\hline
\textbf{Model} & \textbf{LogLikelihood} & \textbf{Parameters} & \textbf{AIC} & \textbf{BIC} & \textbf{AIC Ratio} \\
\hline
ZI-Zipf & -3,102,980 & 2 & 6,205,964 & 6,206,005  & 1.000 \\
ZINB    & -$\infty$  & 3 & $\infty$ & $\infty$    & $\infty$ \\
ZIGP    & NaN        & 3 & NaN      & NaN         & NaN \\
\hline
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\caption{Estimated Parameters for Distribution Models}
\begin{tabular}{lrrr}
\hline
\textbf{Model} & \textbf{Parameter 1} & \textbf{Parameter 2} & \textbf{Parameter 3} \\
\hline
ZIGP & $\eta = 0.9999$ & $\lambda = 46.5641$ & $\theta = 0$ \\
ZI-Zipf & $p_0 = 0.9999348$ & $\alpha = 8.288837$ & -- \\
ZINB & $p_0 = 0.9999348$ & $\pi_0 = 0.9998953$ & size = $13.84857$, $\mu = 1.010229$ \\
\hline
\end{tabular}
  \end{table}

- For ZIGP Model (NaN result):

  -  With eta = 0.9999 (extremely high zero-inflation) and theta = 0.03459262, the ZIGP model is not far from a Zero-Inflated Poisson with probability of zero very close to 1.
  -  When calculating log-likelihood for this model, you're likely getting extremely small probability values for non-zero counts, which when logged are approaching negative infinity.
  -  The combination of multiple very small probabilities in the log-likelihood calculation is causing arithmetic underflow, resulting in NaN.

- For ZINB Model (Inf result):

  -   With extremely high zero-inflation ($p_0 = 0.9999348 and pi_0 = 0.9998953), the model is struggling to explain the non-zero counts.
  -  The parameter size = 13.84857 with mu = 1.010229 means the distribution has a relatively narrow shape.
  -  The log-likelihood is likely resulting in -Inf because some observed counts have essentially zero probability under this model, and log(0) is -Inf.

The Core Issue: Extreme Zero-Inflation

- Your data has extreme zero-inflation (over 99.99% zeros). This causes numerical instabilities when fitting these models:

  -  When eta or p0 are extremely close to 1, the probability assigned to any non-zero count becomes vanishingly small.
  -  This leads to arithmetic underflow, where the computer can't represent such small numbers accurately.
  -  The ZI-Zipf model handles this better because power-law distributions naturally accommodate heavy tails better than Poisson or Negative Binomial.

# Power law

- A true power law would show a straight line. Deviations suggest a different distribution.
```{R, echo=FALSE, eval=FALSE}
# Plot log(frequency) vs log(count)
png("./script3/power_law.png")
plot(log(counts[counts > 0]), log(freqs[counts > 0]), 
     xlab="log(count)", ylab="log(frequency)")
abline(lm(log(freqs[counts > 0]) ~ log(counts[counts > 0])))
dev.off()

```

```{R, powerlaw, echo=FALSE, eval=TRUE, fig.cap="Plot of log(frequency) vs log(count). A power law would show a straight line", out.width = '50%',fig.align = 'center'}
knitr::include_graphics("./script3/power_law.png")
```

# thresholding

```{R, echo= TRUE, eval=FALSE}
# Function to calculate meaningful thresholds for each model
calculate_thresholds <- function(counts, freqs, n_zeros, model_params) {
  # Total count
  n_total <- sum(freqs) + n_zeros

  # Calculate mean and SD of positive counts
  mean_pos <- sum(counts * freqs) / sum(freqs)
  var_pos <- sum(freqs * (counts - mean_pos)^2) / sum(freqs)
  sd_pos <- sqrt(var_pos)

  # Calculate various thresholds
  thresholds <- list()

  # 1. Empirical approaches
  thresholds$empirical <- list(
    mean_plus_2sd = ceiling(mean_pos + 2 * sd_pos),
    mean_plus_3sd = ceiling(mean_pos + 3 * sd_pos),
    mean_plus_5sd = ceiling(mean_pos + 5 * sd_pos),
    percentile_99 = as.numeric(quantile(rep(counts, freqs), 0.99)),
    percentile_999 = as.numeric(quantile(rep(counts, freqs), 0.999))
  )

  # 2. Model-based approaches
  # a) For Zipf
  if (!is.null(model_params$ZI_Zipf)) {
    alpha <- model_params$ZI_Zipf$alpha
    p0 <- model_params$ZI_Zipf$p0

    # Find where probability becomes small
    find_zipf_threshold <- function(p_threshold) {
      count <- 1
      while(count < 100) {  # Reasonable upper limit
        prob <- (1-p0) * zipf_pmf(count, alpha)
        if (prob < p_threshold) break
        count <- count + 1
      }
      return(count)
    }

    thresholds$zipf <- list(
      prob_0001 = find_zipf_threshold(0.0001),
      prob_00001 = find_zipf_threshold(0.00001)
    )
  }

  # b) For ZIGP
  if (!is.null(model_params$ZIGP)) {
    eta <- model_params$ZIGP$eta
    lambda <- model_params$ZIGP$lambda
    theta <- model_params$ZIGP$theta

    # Find where probability becomes small
    find_zigp_threshold <- function(p_threshold) {
      count <- 1
      while(count < 100) {  # Reasonable upper limit
        prob <- (1-eta) * (lambda * (lambda + theta * count)^(count - 1) * 
                 exp(-lambda - theta * count)) / factorial(count)
        if (prob < p_threshold) break
        count <- count + 1
      }
      return(count)
    }

    thresholds$zigp <- list(
      prob_0001 = find_zigp_threshold(0.0001),
      prob_00001 = find_zigp_threshold(0.00001)
    )
  }

  # 3. Count-based approaches
  # Find natural break points in the distribution
  natural_breaks <- find_natural_breaks(counts, freqs)
  thresholds$breaks <- natural_breaks

  # Return all thresholds
  return(thresholds)
}

# Helper function to identify natural break points in count distribution
find_natural_breaks <- function(counts, freqs) {
  # Calculate rate of change between consecutive counts
  rate_change <- numeric(length(counts) - 1)
  for (i in 1:(length(counts)-1)) {
    if (counts[i+1] - counts[i] == 1) {  # Only consider consecutive integers
      rate_change[i] <- freqs[i] / freqs[i+1]
    } else {
      rate_change[i] <- NA  # Mark non-consecutive values
    }
  }

  # Find points where rate of change is significantly higher than average
  valid_rates <- rate_change[!is.na(rate_change)]
  if (length(valid_rates) < 3) return(list(break_points = NULL))

  mean_rate <- mean(valid_rates)
  sd_rate <- sd(valid_rates)

  # Points where rate of change is > mean + 1.5*SD
  break_indices <- which(rate_change > mean_rate + 1.5*sd_rate)
  break_points <- counts[break_indices]

  # Return the identified break points
  return(list(
    break_points = break_points,
    rate_change = rate_change,
    suggested_threshold = if(length(break_points) > 0) min(break_points) + 1 else NULL
  ))
}

thresholds <- calculate_thresholds(counts, freqs, n_zeros, results$parameters)
print(thresholds)
## empirical
## $empirical$mean_plus_2sd
## [1] 4

## $empirical$mean_plus_3sd
## [1] 5

## $empirical$mean_plus_5sd
## [1] 7

## $empirical$percentile_99
## [1] 1

## $empirical$percentile_999
## [1] 2

## $zipf
## $zipf$prob_0001
## [1] 1

## $zipf$prob_00001
## [1] 2

## $zigp
## $zigp$prob_0001
## [1] 1

## $zigp$prob_00001
## [1] 1

## $breaks
## $breaks$break_points
## [1] 1

## $breaks$rate_change
##  [1] 694.980861   4.696629   3.068966   1.318182   1.571429   1.272727   1.571429   1.400000   1.250000   1.333333   1.000000   1.000000   0.750000   4.000000   1.000000   1.000000
## [17]   1.000000         NA   2.000000         NA   1.000000   0.500000         NA         NA         NA         NA   1.000000         NA         NA   0.500000   2.000000         NA
## [33]   1.000000         NA         NA         NA         NA         NA

## $breaks$suggested_threshold
## [1] 2

find_natural_breaks (counts, freqs)
## $break_points
## [1] 1

## $rate_change
##  [1] 694.980861   4.696629   3.068966   1.318182   1.571429   1.272727   1.571429   1.400000   1.250000   1.333333   1.000000   1.000000   0.750000   4.000000   1.000000   1.000000
## [17]   1.000000         NA   2.000000         NA   1.000000   0.500000         NA         NA         NA         NA   1.000000         NA         NA   0.500000   2.000000         NA
## [33]   1.000000         NA         NA         NA         NA         NA

## $suggested_threshold
## [1] 2

```

```{R, echo=FALSE, eval=FALSE}
# more refined threshholding

# Robust Empirical Thresholding for Interaction Counts
# This approach works for both well-fitted distributions and those that don't follow 
# standard theoretical distributions

empirical_threshold <- function(counts, freqs, n_zeros = 0) {
  # Create a data frame for easier analysis
  df <- data.frame(
    count = counts,
    freq = freqs,
    prop = freqs / sum(freqs)
  )

  # Order by count
  df <- df[order(df$count), ]

  # Calculate cumulative proportion (excluding zeros)
  df$cum_prop <- cumsum(df$prop)

  # Calculate several thresholds
  thresholds <- list()

  # 1. Basic statistics on positive counts
  mean_pos <- sum(df$count * df$freq) / sum(df$freq)
  var_pos <- sum(df$freq * (df$count - mean_pos)^2) / sum(df$freq)
  sd_pos <- sqrt(var_pos)

  # 2. Compute empirical percentiles
  weighted_data <- rep(df$count, times = df$freq)

  thresholds$basic <- list(
    mean = mean_pos,
    sd = sd_pos,
    threshold_2sd = ceiling(mean_pos + 2 * sd_pos),
    threshold_3sd = ceiling(mean_pos + 3 * sd_pos),
    p95 = as.numeric(quantile(weighted_data, 0.95)),
    p99 = as.numeric(quantile(weighted_data, 0.99)),
    p999 = as.numeric(quantile(weighted_data, 0.999))
  )

  # 3. Detect "elbows" or breakpoints in the distribution
  # Calculate rate of change in frequency between consecutive counts
  df$rate_change <- NA
  for (i in 1:(nrow(df)-1)) {
    if (df$count[i+1] - df$count[i] == 1) {  # Only for consecutive counts
      df$rate_change[i] <- df$freq[i] / df$freq[i+1]
    }
  }

  valid_rates <- df$rate_change[!is.na(df$rate_change)]
  if (length(valid_rates) >= 3) {
    mean_rate <- mean(valid_rates)
    sd_rate <- sd(valid_rates)

    # Find significant jumps in rate of change
    # Points where rate of change is > mean + 1.5*SD
    break_indices <- which(df$rate_change > mean_rate + 1.5*sd_rate)
    break_points <- df$count[break_indices]

    thresholds$breakpoints <- list(
      points = break_points,
      suggested = if(length(break_points) > 0) min(break_points) + 1 else NULL
    )
  } else {
    thresholds$breakpoints <- list(
      points = NULL,
      suggested = NULL
    )
  }

  # 4. Entropy-based thresholding
  # Calculate entropy of frequency distribution
  p <- df$prop
  entropy <- -sum(p * log(p))

  # Find threshold that maximizes entropy difference between below and above threshold
  max_diff <- -Inf
  best_threshold <- 1

  for (i in 1:max(df$count)) {
    below <- df[df$count <= i, ]
    above <- df[df$count > i, ]

    # Skip if either group is empty
    if (nrow(below) == 0 || nrow(above) == 0) next

    # Normalize probabilities within each group
    p_below <- below$freq / sum(below$freq)
    p_above <- above$freq / sum(above$freq)

    # Calculate entropies
    entropy_below <- -sum(p_below * log(p_below))
    entropy_above <- -sum(p_above * log(p_above))

    # Calculate weighted entropy difference
    entropy_diff <- (sum(below$freq) * entropy_below + sum(above$freq) * entropy_above) / sum(df$freq)

    if (entropy - entropy_diff > max_diff) {
      max_diff <- entropy - entropy_diff
      best_threshold <- i
    }
  }

  thresholds$entropy <- list(
    threshold = best_threshold + 1  # Add 1 for exclusive threshold
  )

  # 5. Exponential decay test
  # Test if counts follow exponential decay by fitting log(frequency) ~ count
  if (nrow(df) > 5) {
    log_freq <- log(df$freq)
    model <- lm(log_freq ~ df$count)
    r_squared <- summary(model)$r.squared

    # If strong exponential decay, use where predicted frequency < 1
    if (r_squared > 0.8) {
      coefs <- coef(model)
      threshold_exp <- ceiling(-coefs[1] / coefs[2])  # Where predicted freq < 1
      thresholds$exponential <- list(
        r_squared = r_squared,
        threshold = threshold_exp
      )
    } else {
      thresholds$exponential <- list(
        r_squared = r_squared,
        threshold = NULL
      )
    }
  }

  # 6. Rarity-based threshold
  # Define thresholds based on observation frequency
  total_observations <- sum(df$freq)
  rare_threshold <- ceiling(total_observations * 0.001)  # 0.1% of observations

  rare_counts <- df[df$freq <= rare_threshold, "count"]
  if (length(rare_counts) > 0) {
    thresholds$rarity <- list(
      threshold = min(rare_counts)
    )
  } else {
    thresholds$rarity <- list(
      threshold = NULL
    )
  }

  # 7. Add zero-inflation context if provided
  if (n_zeros > 0) {
    # Calculate proportion of zeros
    p_zero <- n_zeros / (n_zeros + sum(df$freq))
    thresholds$zero_inflation <- list(
      p_zero = p_zero
    )

    # For extreme zero-inflation (>99.9%), more aggressive thresholds
    if (p_zero > 0.999) {
      # With extreme zeros, any occurrence might be significant
      thresholds$extreme_zero_inflation <- list(
        threshold = 2  # Consider anything geq 2 significant when zeros dominate
      )
    }
  }

  # 8. Final recommended threshold based on all methods
  # Logic for combining thresholds
  candidates <- c(
    thresholds$basic$threshold_2sd,
    thresholds$breakpoints$suggested,
    thresholds$entropy$threshold,
    thresholds$rarity$threshold
  )

  # Add exponential threshold if available
  if (!is.null(thresholds$exponential$threshold)) {
    candidates <- c(candidates, thresholds$exponential$threshold)
  }

  # Add extreme zero-inflation threshold if applicable
  if (!is.null(thresholds$extreme_zero_inflation$threshold)) {
    candidates <- c(candidates, thresholds$extreme_zero_inflation$threshold)
  }

  # Remove NAs
  candidates <- candidates[!is.na(candidates)]

  # For final recommendation, use median of available thresholds
  if (length(candidates) > 0) {
    thresholds$recommended <- median(candidates, na.rm = TRUE)
  } else {
    # Default if no thresholds available
    thresholds$recommended <- max(2, ceiling(mean_pos + 2 * sd_pos))
  }

  # Provide both results and visualization
  results <- list(
    thresholds = thresholds,
    data = df
  )

  # Return everything
  return(results)
}

# Function to visualize the thresholds on a count distribution
plot_thresholds <- function(threshold_results, log_scale = TRUE) {
  df <- threshold_results$data
  thresholds <- threshold_results$thresholds

  # Setup plot
  par(mfrow = c(1, 1))

  # Create bar plot of frequencies
  if (log_scale) {
    barplot(log(df$freq), names.arg = df$count, xlab = "Count", ylab = "Log Frequency", 
            main = "Distribution of Counts with Thresholds", col = "lightblue")
  } else {
    barplot(df$freq, names.arg = df$count, xlab = "Count", ylab = "Frequency", 
            main = "Distribution of Counts with Thresholds", col = "lightblue")
  }

  # Add recommended threshold line
  abline(v = which(df$count == ceiling(thresholds$recommended)), col = "red", lwd = 2)

  # Add other threshold lines
  # Add 2SD threshold
  if (!is.null(thresholds$basic$threshold_2sd)) {
    pos <- which(df$count >= thresholds$basic$threshold_2sd)[1]
    if (!is.na(pos)) {
      abline(v = pos, col = "blue", lty = 2)
    }
  }

  # Add breakpoint threshold
  if (!is.null(thresholds$breakpoints$suggested)) {
    pos <- which(df$count >= thresholds$breakpoints$suggested)[1]
    if (!is.na(pos)) {
      abline(v = pos, col = "green", lty = 2)
    }
  }

  # Add entropy threshold
  if (!is.null(thresholds$entropy$threshold)) {
    pos <- which(df$count >= thresholds$entropy$threshold)[1]
    if (!is.na(pos)) {
      abline(v = pos, col = "purple", lty = 2)
    }
  }

  # Add legend
  legend("topright", 
         legend = c("Recommended", "2SD", "Breakpoint", "Entropy"),
         col = c("red", "blue", "green", "purple"),
         lty = c(1, 2, 2, 2),
         lwd = c(2, 1, 1, 1))

  # Return the plot
  invisible(threshold_results)
}

rheumatoid_thresholds <- empirical_threshold(counts, freqs, n_zeros)
names(rheumatoid_thresholds$thresholds)
#[1] "basic"                  "breakpoints"            "entropy"                "exponential"            "rarity"                 "zero_inflation"         "extreme_zero_inflation"
#[8] "recommended"           
rheumatoid_thresholds$thresholds$basic$p95$threshold_2sd
rheumatoid_thresholds$thresholds$basic$threshold_2sd
rheumatoid_thresholds$thresholds$breakpoints$suggested
rheumatoid_thresholds$thresholds$entropy
rheumatoid_thresholds$thresholds$exponential$threshold
rheumatoid_thresholds$thresholds$rarity$threshold
rheumatoid_thresholds$thresholds$zero_inflation
rheumatoid_thresholds$thresholds$extreme_zero_inflation$threshold
rheumatoid_thresholds$thresholds$recommended

plot_thresholds(rheumatoid_thresholds)
print(rheumatoid_thresholds$thresholds$recommended)

```

# plink preprocessing

```{R, echo=FALSE, eval=FALSE}
maps <- read.table("Cleaned_CaseControl_RAauto_pruned200kb_.3.bim",header=FALSE)
#.bim (PLINK extended MAP file)
#Extended variant information file accompanying a .bed binary genotype table. (--make-just-bim can be used to update just this file.)
#A text file with no header line, and one line per variant with the following six fields:
#    Chromosome code (either an integer, or 'X'/'Y'/'XY'/'MT'; '0' indicates unknown) or name
#    Variant identifier
#    Position in morgans or centimorgans (safe to use dummy value of '0')
#    Base-pair coordinate (1-based; limited to 231-2)
#    Allele 1 (corresponding to clear bits in .bed; usually minor)
#    Allele 2 (corresponding to set bits in .bed; usually major)

## .fam (PLINK sample information file)
## Sample information file accompanying a .bed binary genotype table. (--make-just-fam can be used to update just this file.) Also generated by "--recode lgen" and "--recode rlist".
## A text file with no header line, and one line per sample with the following six fields:
##     Family ID ('FID')
##     Within-family ID ('IID'; cannot be '0')
##     Within-family ID of father ('0' if father isn't in dataset)
##     Within-family ID of mother ('0' if mother isn't in dataset)
##     Sex code ('1' = male, '2' = female, '0' = unknown)
##     Phenotype value ('1' = control, '2' = case, '-9'/'0'/non-numeric = missing data if case/control)
## With the use of additional loading flag(s), PLINK can also correctly interpret some .fam files missing one or more of these fields.

maps$chrNpos <- paste0(maps$V1,"_",maps$V4)
lm_results$SNP1ID <- ""
lm_results$SNP2ID <- ""

for(s in 1:length(SNPs)){
  tmp1 <- strsplit(SNPs[s]," ")[[1]][1]
  tmp2 <- strsplit(SNPs[s]," ")[[1]][2]

  index <- which(maps$chrNpos==tmp1)
  lm_results$SNP1ID[s] <- as.character(maps[index,2])
  index <- which(maps$chrNpos==tmp2)
  lm_results$SNP2ID[s] <- as.character(maps[index,2])

}

index <- which(lm_results$adj_p_value<0.05)
allSNPs <- union(lm_results$SNP1ID,lm_results$SNP2ID)

for(p in 1:length(index)){
  plinkCommand <- paste0("plink --bfile Cleaned_CaseControl_RAauto_pruned200kb_.3 --ld   ",lm_results$SNP1ID[p]," " ,lm_results$SNP2ID[p]," --out  ",lm_results$SNP1ID[p],"_" ,lm_results$SNP2ID[p])
  system(plinkCommand)
}

## PLINK v1.90p 64-bit (30 Nov 2019)              www.cog-genomics.org/plink/1.9/
## (C) 2005-2019 Shaun Purcell, Christopher Chang   GNU General Public License v3
## Logging to rs3134926_rs9268831.log.
## Options in effect:
##   --bfile Cleaned_CaseControl_RAauto_pruned200kb_.3
##   --ld rs3134926 rs9268831
##   --out rs3134926_rs9268831

## 515398 MB RAM detected; reserving 257699 MB for main workspace.
## Error: Failed to open Cleaned_CaseControl_RAauto_pruned200kb_.3.bed.
## PLINK v1.90p 64-bit (30 Nov 2019)              www.cog-genomics.org/plink/1.9/
## (C) 2005-2019 Shaun Purcell, Christopher Chang   GNU General Public License v3
## Logging to _.log.
## Options in effect:
##   --bfile Cleaned_CaseControl_RAauto_pruned200kb_.3
##   --ld
##   --out _

## Error: --ld requires at least 2 parameters.
## For more information, try "plink --help <flag name>" or "plink --help | more".

lm_results$LD <- ""
lm_results$Rsq <- ""
lm_results$Dprime <- ""
library(reader)
for(r in 1:length(index)){
  file <- paste0(lm_results$SNP1ID[r],"_",lm_results$SNP2ID[r],".log")
  lds <- n.readLines(file,n=24, skip = 23)[1]
  tmpR <- strsplit(lds,"D'")[[1]][1]
  tmpD <- strsplit(lds,"D'")[[1]][2]
  lm_results$Rsq[r] <- as.numeric(strsplit(tmpR,"R-sq = ")[[1]][2])
  lm_results$Dprime[r] <- as.numeric(strsplit(tmpD," <- ")[[1]][2])
  lm_results$LD[r]=lds
}

write.csv(lm_results[1:length(index),],paste0(dirOut,"/interacting_SNPs.csv"))

# counts=countMatrix#readRDS(paste0(dirOut,"/countMatrix_Nit",t,"_",".rds"))
# rm(countMatrix)
# gc()
# 
# counts[lower.tri(counts, diag = T)] =0# counts[upper.tri(counts)]
# index = which(counts>0)
# rowIndex = index%%ncol(counts)
# colIndex = floor(index/ncol(counts))+1
# length(index)
# 
# interactions = data.frame(SNP1=colnames(counts)[rowIndex],SNP2=colnames(counts)[colIndex],Index=index,counts=counts[index])
# interactions = interactions[order(interactions$counts,decreasing = T),]
# write.csv(interactions,paste0(dirOut,"/interactions",t,".csv"),row.names = F)
# 
# ######################## Interaction modelling and code here
# #interactions=read.csv(paste0(dirOut,"/interactions",t,".csv"))
# index=which(interactions$counts>=2)
# 
# lm_results = data.frame(interactions[index,],SNP1_effect=0,SNP1_Pvalue=1,SNP2_effect=0,
#                         SNP2_Pvalue=1,int_effect=0,int_Pvalue=1)
# dim(lm_results)
# 
# for(l in 1:nrow(lm_results)){
#   ind1 = which(colnames(dt)==lm_results$SNP1[l])
#   ind2 = which(colnames(dt)==lm_results$SNP2[l])
#   s = summary(lm(dt$pheno ~ dt[,ind1]*dt[,ind2]), data= dt)$coeff
#   if(nrow(s)==4){
#     lm_results$SNP1_effect[l] = s[2,1]
#     lm_results$SNP1_Pvalue[l] = s[2,4]
#     lm_results$SNP2_effect[l] = s[3,1]
#     lm_results$SNP2_Pvalue[l] = s[3,4]
#     lm_results$int_effect[l] = s[4,1]
#     lm_results$int_Pvalue[l] = s[4,4]
#   }
# }
# 
# 
# 
# lm_results = lm_results[order(lm_results$int_Pvalue,decreasing = F),]
# lm_results$adj_p = lm_results$int_Pvalue*(ncol(dt)-2)*(ncol(dt)-3)*.5
# # lm_results$adj_int_Pvalue = lm_results$int_Pvalue*(0.5*ncol(geno)*(ncol(geno)-1))
# write.csv(lm_results,paste0(dirOut,"/lm_results",t,".csv"),row.names = F)

```