### 2025-05-04 cover_letter.tex
smale changes to paper

\documentclass{article}

\usepackage{charter} % Use the Charter font for the document text

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr,graphicx}

\fancypagestyle{firstpage}{%
  \fancyhf{}% Clear header/footer
  \renewcommand{\headrulewidth}{0pt}%
  \renewcommand{\footrulewidth}{1pt}%
  % Add more detail here if needed
}
\fancypagestyle{otherpages}{%
  \fancyhf{}% Clear header/footer
  \renewcommand{\headrulewidth}{1pt}%
  \renewcommand{\footrulewidth}{1pt}%
  % Add more detail here if needed
}

\AtBeginDocument{\thispagestyle{firstpage}}
\pagestyle{otherpages}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex}

\begin{document}

\vspace*{\dimexpr-\headsep-\headheight-1pt}

\noindent
\begin{minipage}[t]{0.5\textwidth}
  \includegraphics[width=4cm]{CSIRO_logo.png}
\end{minipage}%
\begin{minipage}[t]{0.5\textwidth}
  \raggedleft
  \raisebox{1cm}{\begin{tabular}{ l @{} }
  \today \\[12pt] % Date
   Robert Dunne \\
   13 Garden Street\\
    Eveleigh NSW \\
    Phone: (+61) 439 905  075 \\
    Email: rob.dunne@csiro.au
  \end{tabular}}
\end{minipage}

\rule{\linewidth}{1pt}

\bigskip\bigskip

Dr. Christiana Gousiadi

Research Networks AS, Oslo, Norway

\noindent
\begin{minipage}[t]{\textwidth}
  Dr. Christiana Gousiadi \\ 
  Research Networks AS, Oslo, Norway \\
  Editor-in-Chief,  Computational and Structural Biotechnology Reports \\ 
\end{minipage}

\bigskip\bigskip

% Start of the letter content
\noindent
Dear Dr. Christiana Gousiadi,

\bigskip

I am sending you the manuscript of our research paper entitled ``Detection of Ecstatic Interactions in High-Dimensional
Genomic Data Using Random Forests''. We would like to have the manuscript considered for publication in  Computational and Structural Biotechnology Reports.

Epistatic interactions can play an important role in the genetic mechanisms that control phenotypic variation. However,
identifying these interactions in high dimensional genomic data can be very challenging due to the large computational
burden.

Random Forests Decision Trees are widely used in a variety of disciplines and handle interactions well.
In fact, in a Random Forest model there are no ``main effects'', they are all interactions.

However, Random Forests models do not explicitly detect variable interactions. We have addressed this problem with a
computationally feasible approach.

We confirm that this manuscript has not been published elsewhere and is not under consideration by another journal.
All authors have approved the manuscript and agree with its submission to [insert the name of the target journal].

We believe that this paper is well aligned with the aims and scope of the journal Computational and Structural
Biotechnology Reports.

Robert Dunne

\end{document}