### 2025-07-02 Supplementary_Files/Supplementary_File_Comparison_with_BOOST.Rmd
added Supplementary_Files and bib items

---
title: "Comparison with BOOST"
author: "Rob Dunne"
date: "Monday, June 16, 2025"
output: 
    bookdown::pdf_document2:
     toc: true
     toc_depth: 3
     fig_caption: yes
     citation_package: biblatex
bibliography: 
    ../epistaticInteractions.bib 
header-includes:
  \usepackage{dcolumn}
  \usepackage{placeins}
  \usepackage{float}
  \usepackage{biblatex}
  \usepackage{caption}
  \usepackage{subcaption}
---

# load data

- 4798 samples
- 94485 SNPs

```{r, echo=FALSE, eval=FALSE}
library(lmtest)
library(ranger)
#library(RCy3)

#Noise <- 0.1
numIte  <-  1
t <- numIte
ntree <- 1000

dirOut  <-  paste0("Results")
if(!dir.exists(dirOut)){
  dir.create(dirOut)
}
#dirOut  <-  paste0("Results/pruned200K.3r2_geno_Missing_Averaged")
dt  <-  readRDS("pruned200K.3r2_geno_Missing_Averaged.rds")
pheno <- read.table("Cleaned_CaseControl_RAauto.fam",header=F)
pheno$ID <- paste0(pheno$V1,"_",pheno$V2)
all.equal(row.names(dt),pheno$ID)
dt <- data.frame(pheno=as.factor(pheno$V6),dt)
dim(dt) #[1]  4798 94485

#94,486 SNPs (4.46 billion possible pairwise interaction tests) and 4,798 individuals.

dt$pheno <- as.factor(dt$pheno)
#dt <- dt[which(dt$treatment=="short"),-2]

# quick load of results etc
#dirOut  <-  paste0("Results")
#load(file="Results/var_importance_gini.Rdata")
#load(file=paste0(dirOut,"/All_Interactions.Rdata"))
#load(file="./Results/interaction_snps_counts_fixed.Rdata")
#load(file=paste0(dirOut,"/lm_results_fixed.Rdata"))

```

# BOOST

To demonstrate the computational advantage of our Random Forest approach, we attempted exhaustive pairwise epistasis
testing using PLINK v1.9 on the full rheumatoid arthritis dataset (94,484 SNPs, requiring 4.46 billion tests). Despite
allocating 257GB of RAM, the analysis resulted in segmentation faults due to computational and memory limitations. This
demonstrates the practical impossibility of exhaustive approaches on genome-wide datasets.

In contrast, our Random Forest method completed analysis of the same dataset in [X] minutes, identifying [Y] candidate
interactions for follow-up testing. To enable direct comparison, we implemented Boolean operation screening (similar to
BOOST) on random subsets of 50,000 SNP pairs, showing [Z]% overlap in top-ranked interactions while requiring [W]-fold
less computation time.

We have used the BOOST (BOOlean operation-based Screening and Testing) @wanBOOSTFastApproach2010 as a comparitor.

- Stage 1 (Screening): Uses Boolean operations to quickly identify potentially interacting SNP pairs

    - For each SNP pair, creates a 3×3 contingency table (0,1,2 genotypes for each SNP)
    -    Applies Boolean functions (AND, OR, XOR) to identify patterns
    -    Ranks pairs by how well Boolean operations separate cases from controls
- Stage 2 (Testing): Applies statistical tests to top-ranked pairs

   - Uses logistic regression or other statistical tests
   -  Applies multiple testing corrections

However, even with over 500GB of RAM, the algorithm could not run on this data set.
We implemented the core BOOT functionality in R and and ran it on 500000 randomly selected SNPS.
To process the whole data set would have taken over 4 days with 64 cores.

```{R, echo=FALSE, eval=FALSE}
# Convert your data to PLINK format
library(snpStats)
# Assuming your data is in the format you used for RF
temp<-as.matrix(dt[,-1])
ptemp<- dt[,1]
temp_snpmatrix <- temp + 1  # Convert 0,1,2 to 1,2,3
temp_snpmatrix[is.na(temp)] <- 0  # Set missing to 0

# Create SnpMatrix object
snp_matrix <- new("SnpMatrix", temp_snpmatrix)

write.plink("boost_input", snps =snp_matrix,
            phenotype = ptemp, 
            id = rownames(dt)
            )
#Writing FAM file to boost_input.fam 
#Writing extended MAP file to boost_input.bim 
#Writing BED file to boost_input.bed (SNP-major mode)
#NULL

```
```{verbatim}
> plink --bfile boost_input --epistasis --allow-no-sex --out epistasis_results
PLINK v1.90p 64-bit (30 Nov 2019)              www.cog-genomics.org/plink/1.9/
(C) 2005-2019 Shaun Purcell, Christopher Chang   GNU General Public License v3
Logging to epistasis_results.log.
Options in effect:
  --allow-no-sex
  --bfile boost_input
  --epistasis
  --out epistasis_results
515555 MB RAM detected; reserving 257777 MB for main workspace.
94484 variants loaded from .bim file.
4798 people (0 males, 0 females, 4798 ambiguous) loaded from .fam.
Ambiguous sex IDs written to epistasis_results.nosex .
4798 phenotype values loaded from .fam.
Using up to 63 threads (change this with --threads).
Before main variant filters, 4798 founders and 0 nonfounders present.
Calculating allele frequencies... 0%Segmentation fault (core dumped)
```

## Boost in R

```{R, echo=FALSE, eval=FALSE}
# Use shared memory approach (no data copying)
library(parallel)

# This approach doesn't copy data to each worker
boolean_shared_memory <- function(snp_data, phenotype, max_pairs = 100000, n_cores = 20) {

  # Use mclapply instead of makeCluster (shares memory)
  n_snps <- ncol(snp_data)

  # Generate random pairs
  set.seed(123)
  random_pairs <- replicate(max_pairs, {
    i <- sample(1:(n_snps-1), 1)
    j <- sample((i+1):n_snps, 1)
    c(i, j)
  }, simplify = FALSE)

  # Split into chunks
  chunk_size <- ceiling(length(random_pairs) / n_cores)
  chunks <- split(random_pairs, ceiling(seq_along(random_pairs) / chunk_size))

  cat("Using mclapply with", n_cores, "cores (shared memory)\n")

  # mclapply shares memory instead of copying
  results_list <- mclapply(chunks, function(chunk) {
    chunk_results <- data.frame()

    for(pair in chunk) {
      i <- pair[1]
      j <- pair[2]

      snp1 <- snp_data[, i]  # References original data, no copy
      snp2 <- snp_data[, j]

      if (length(unique(snp1)) < 2 || length(unique(snp2)) < 2) next

      tryCatch({
        xor_op <- (snp1 >= 1) != (snp2 >= 1)
        test_result <- fisher.test(table(xor_op, phenotype))

        chunk_results <- rbind(chunk_results, data.frame(
          SNP1_idx = i, SNP2_idx = j,
          p_value = test_result$p.value,
          odds_ratio = as.numeric(test_result$estimate)
        ))
      }, error = function(e) {})
    }

    return(chunk_results)
  }, mc.cores = n_cores)

  final_results <- do.call(rbind, results_list)
  return(final_results[order(final_results$p_value), ])
}
# This should use much less memory since it doesn't copy data to workers
cat("Trying shared memory approach...\n")

boost_time_shared <- system.time({
  boost_results_shared <- boolean_shared_memory(
    dt[,-1],           # SNP data  
    dt[,1],            # phenotype
    max_pairs = 500000,  # Start with 500k pairs
    n_cores = 60       # Conservative core count
  )
})

# max_pairs = 10000000 failed -- many processes in "D" state, I eventually killed it after about 1 hour
# max_pairs = 5000000 also  failed -- many processes in "D" state, I eventually killed it

cat("Shared memory approach completed in", boost_time_shared[3], "seconds\n")
cat("Found", nrow(boost_results_shared), "tested pairs\n")
print(head(boost_results_shared, 5))

#save(boost_results_shared, file="./script3/boost_results_shared_500000.Rdata")
#load(file="./script3/boost_results_shared_500000.Rdata")

```

```{R, echo=TRUE, eval=FALSE}
boost_time_shared 
#    user   system  elapsed 
#1593.886   99.380   42.000

((choose(94485,2)/500000)*42)/(60*60*24)
#[1] 4.33967 days run time to process the whole data set

print(head(boost_results_shared, 5))
#        SNP1_idx SNP2_idx      p_value odds_ratio
#5.3801     34461    75705 1.031365e-33  0.4767794
#21.4869    34461    58455 4.896431e-32  0.4869738
#6.8174     34461    67197 2.060349e-27  0.5194040
#3.15611    34457    58152 3.514479e-17  0.6043620
#19.6946    17151    34442 9.914728e-15  1.5910823

```
# comparison of RF method and BOOST

We load the **interaction_snps_counts** (calculated in Supp1.Rmd),

```{R, echo=TRUE, eval=FALSE}
load(file="./Results/interaction_snps_counts_fixed.Rdata")
head(interaction_snps_counts)
#              interaction_snps Freq
#281122 X6_32471505 X6_32771829  466
#281468 X6_32682149 X6_32771829  144
#280884 X6_32222493 X6_32771829  113

```

```{R, echo=FALSE, eval=FALSE}

load(file="./script3/boost_results_shared_500000.Rdata")
load(file="./Results/interaction_snps_counts_fixed.Rdata")

# Function to compare RF and BOOST results
compare_rf_boost_results <- function(rf_results, boost_results, snp_names) {

  # Convert BOOST indices to SNP names
  boost_with_names <- boost_results
  boost_with_names$SNP1_name <- snp_names[boost_results$SNP1_idx]
  boost_with_names$SNP2_name <- snp_names[boost_results$SNP2_idx]

  # Create standardized pair identifiers (alphabetical order)
  boost_with_names$pair_id <- apply(boost_with_names[, c("SNP1_name", "SNP2_name")], 1, 
                                   function(x) paste(sort(x), collapse = " "))

  # Process RF results - split the interaction_snps column
  rf_processed <- rf_results
  rf_processed$pair_id <- rf_processed$nteraction_snps

  # Find overlapping pairs
  common_pairs <- intersect(rf_processed$pair_id, boost_with_names$pair_id)

  # Create comparison dataframe
  comparison_results <- data.frame()

  for(pair in common_pairs) {
    rf_row <- rf_processed[rf_processed$pair_id == pair, ]
    boost_row <- boost_with_names[boost_with_names$pair_id == pair, ]

    comparison_results <- rbind(comparison_results, data.frame(
      pair_id = pair,
      rf_frequency = rf_row$Freq[1],
      boost_p_value = boost_row$p_value[1],
      boost_odds_ratio = boost_row$odds_ratio[1],
      rf_rank = which(rf_processed$pair_id == pair)[1],
      boost_rank = which(boost_with_names$pair_id == pair)[1]
    ))
  }

  return(list(
    comparison = comparison_results,
    overlap_count = length(common_pairs),
    total_rf = nrow(rf_processed),
    total_boost = nrow(boost_with_names),
    overlap_percent = length(common_pairs) / min(nrow(rf_processed), nrow(boost_with_names)) * 100
  ))
}

# Apply the comparison
snp_names <- colnames(dt[,-1])  # Get SNP names from your data

comparison_result <- compare_rf_boost_results(
  interaction_snps_counts,  # Your RF results
  boost_results_shared,     # Your BOOST results  
  snp_names
)

# View results
cat("Overlap summary:\n")
cat("Common pairs found:", comparison_result$overlap_count, "\n") #Common pairs found: 28 
cat("RF total pairs:", comparison_result$total_rf, "\n")         #RF total pairs: 291142 
cat("BOOST total pairs:", comparison_result$total_boost, "\n")   #BOOST total pairs: 500000 
cat("Overlap percentage:", round(comparison_result$overlap_percent, 2), "%\n") #Overlap percentage: 0.01 %

print("Top overlapping pairs:")
print(head(comparison_result$comparison[order(comparison_result$comparison$rf_frequency, decreasing = TRUE), ], 10))

# Analyze correlation between RF frequency and BOOST significance
 if(nrow(comparison_result$comparison) > 0) {

  # Correlation between RF frequency and BOOST -log10(p-value)
  rf_freq <- comparison_result$comparison$rf_frequency
  boost_neg_log_p <- -log10(comparison_result$comparison$boost_p_value)

  correlation <- cor(rf_freq, boost_neg_log_p, use = "complete.obs")

  cat("Correlation between RF frequency and BOOST -log10(p-value):", round(correlation, 3), "\n")

  # Plot the relationship
  plot(rf_freq, boost_neg_log_p, 
       xlab = "RF Co-occurrence Frequency", 
       ylab = "BOOST -log10(p-value)",
       main = "RF vs BOOST: Method Comparison",
       pch = 19, col = "blue")

  # Add trend line
  if(!is.na(correlation) && abs(correlation) > 0.1) {
    abline(lm(boost_neg_log_p ~ rf_freq), col = "red", lwd = 2)
    legend("topright", paste("r =", round(correlation, 3)), bty = "n")
  }
}
# very little correlation

# Compare top-ranked pairs from each method
top_rf <- head(interaction_snps_counts[order(interaction_snps_counts$Freq, decreasing = TRUE), ], 20)
top_boost <- head(boost_results_shared[order(boost_results_shared$p_value), ], 20)

cat("Top 20 RF pairs:\n")
print(top_rf)

cat("\nTop 20 BOOST pairs (with names):\n")
top_boost_named <- top_boost
top_boost_named$SNP1_name <- snp_names[top_boost$SNP1_idx]
top_boost_named$SNP2_name <- snp_names[top_boost$SNP2_idx]
print(top_boost_named[, c("SNP1_name", "SNP2_name", "p_value", "odds_ratio")])

# Analyze what types of interactions each method finds
cat("Analysis of method differences:\n")
cat("RF finds", nrow(interaction_snps_counts), "co-occurrence based interactions\n")   #RF finds 291142 co-occurrence based interactions
cat("BOOST finds", sum(boost_results_shared$p_value < 0.05), "Boolean logic interactions\n") #BOOST finds 25436 Boolean logic interactions
cat("Methods are detecting different interaction mechanisms\n")

# Convert BOOST indices to SNP names
boost_with_names <-  boost_results_shared
boost_with_names$SNP1_name <- snp_names[boost_results_shared$SNP1_idx]
boost_with_names$SNP2_name <- snp_names[boost_results_shared$SNP2_idx]

  # Create standardized pair identifiers (alphabetical order)
boost_with_names$pair_id <- apply(boost_with_names[, c("SNP1_name", "SNP2_name")], 1, 
                                   function(x) paste(sort(x), collapse = " "))

# Look at the p-value distribution of overlapping vs non-overlapping
overlapping_pairs <- comparison_result$comparison$pair_id
length(overlapping_pairs) #[1] 28
boost_overlap <- boost_with_names$pair_id %in% overlapping_pairs

boost_non_overlap <- !boost_overlap

cat("BOOST p-values for overlapping pairs:\n")
summary(-log10(boost_results_shared$p_value[boost_overlap]))
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 0.0000  0.3034  0.4397  0.7352  0.8353  4.7758

cat("BOOST p-values for non-overlapping pairs:\n") 
summary(-log10(boost_results_shared$p_value[boost_non_overlap]))
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 0.0000  0.1178  0.2972  0.4342  0.6056 32.9866

```

In the paper we only consider the  640 interaction_snps_counts with Freq > 1

```{R, echo=FALSE, eval=FALSE}
# Add FDR correction and pair_id to boost_results_shared
library(stats)

# Add FDR corrected p-values
boost_results_shared$fdr_p_value <- p.adjust(boost_results_shared$p_value, method = "fdr")

# Add pair_id column
snp_names <- colnames(dt[,-1])
boost_results_shared$SNP1_name <- snp_names[boost_results_shared$SNP1_idx]
boost_results_shared$SNP2_name <- snp_names[boost_results_shared$SNP2_idx]

# Create standardized pair_id (alphabetical order)
boost_results_shared$pair_id <- apply(boost_results_shared[, c("SNP1_name", "SNP2_name")], 1, 
                                     function(x) paste(sort(x), collapse = " "))

# Check the results
cat("BOOST results with FDR correction:\n")
cat("Total pairs tested:", nrow(boost_results_shared), "\n") #500000 
cat("Significant at p < 0.05:", sum(boost_results_shared$p_value < 0.05), "\n")  # 25436 
cat("Significant at FDR < 0.05:", sum(boost_results_shared$fdr_p_value < 0.05), "\n") #55 
cat("Significant at FDR < 0.10:", sum(boost_results_shared$fdr_p_value < 0.10), "\n") #60

print(head(boost_results_shared[, c("pair_id", "p_value", "fdr_p_value", "odds_ratio")], 10))

#2) Focused Comparison: RF Freq > 1 vs Significant BOOST
# Filter to meaningful comparisons
rf_significant <- interaction_snps_counts[interaction_snps_counts$Freq > 1, ]
boost_significant <- boost_results_shared[boost_results_shared$fdr_p_value < 0.10, ]  # Adjust threshold as needed

cat("Focused comparison:\n")
cat("RF pairs with Freq > 1:", nrow(rf_significant), "\n")              #640 
cat("BOOST pairs with FDR < 0.10:", nrow(boost_significant), "\n")      #60

# Updated comparison function
compare_rf_boost_focused <- function(rf_results, boost_results) {

  # Find overlapping pairs
  common_pairs <- intersect(rf_results$interaction_snps, boost_results$pair_id)

  # Create detailed comparison for overlapping pairs
  comparison_detailed <- data.frame()

  for(pair in common_pairs) {
    rf_row <- rf_results[rf_results$interaction_snps == pair, ]
    boost_row <- boost_results[boost_results$pair_id == pair, ]

    comparison_detailed <- rbind(comparison_detailed, data.frame(
      pair_id = pair,
      rf_frequency = rf_row$Freq[1],
      boost_p_value = boost_row$p_value[1],
      boost_fdr_p_value = boost_row$fdr_p_value[1],
      boost_odds_ratio = boost_row$odds_ratio[1],
      rf_rank = which(rf_results$interaction_snps == pair)[1],
      boost_rank = which(boost_results$pair_id == pair)[1]
    ))
  }

  # Summary statistics
  overlap_stats <- list(
    total_overlap = length(common_pairs),
    rf_total = nrow(rf_results),
    boost_total = nrow(boost_results), 
    overlap_percent_rf = length(common_pairs) / nrow(rf_results) * 100,
    overlap_percent_boost = length(common_pairs) / nrow(boost_results) * 100,
    comparison_table = comparison_detailed
  )

  return(overlap_stats)
}

# Perform focused comparison
focused_comparison <- compare_rf_boost_focused(rf_significant, boost_significant)

cat("\n=== FOCUSED COMPARISON RESULTS ===\n")
cat("RF significant pairs (Freq > 1):", focused_comparison$rf_total, "\n")                         #640 
cat("BOOST significant pairs (FDR < 0.10):", focused_comparison$boost_total, "\n")                 #60 
cat("Overlapping pairs:", focused_comparison$total_overlap, "\n")                                  #Overlapping pairs: 0 
cat("Overlap as % of RF significant:", round(focused_comparison$overlap_percent_rf, 2), "%\n")
cat("Overlap as % of BOOST significant:", round(focused_comparison$overlap_percent_boost, 2), "%\n")

if(nrow(focused_comparison$comparison_table) > 0) {
  cat("\nOverlapping pairs details:\n")
  print(focused_comparison$comparison_table[order(focused_comparison$comparison_table$rf_frequency, decreasing = TRUE), ])
} else {
  cat("\nNo overlapping pairs found between RF (Freq > 1) and BOOST (FDR < 0.10)\n")
}
#No overlapping pairs found between RF (Freq > 1) and BOOST (FDR < 0.10)

#3) Alternative Comparison Strategy
#If there's still little overlap, let's try a more lenient approach:

# More lenient comparison
lenient_comparison <- function() {

  # Use top performers from each method
  top_rf <- head(rf_significant[order(rf_significant$Freq, decreasing = TRUE), ], 100)
  top_boost <- head(boost_results_shared[order(boost_results_shared$p_value), ], 100)

  common_pairs_lenient <- intersect(top_rf$interaction_snps, top_boost$pair_id)

  cat("=== LENIENT COMPARISON (Top 100 each) ===\n")
  cat("Top 100 RF pairs overlap with top 100 BOOST pairs:", length(common_pairs_lenient), "\n")

  if(length(common_pairs_lenient) > 0) {
    for(pair in common_pairs_lenient) {
      rf_info <- top_rf[top_rf$interaction_snps == pair, ]
      boost_info <- top_boost[top_boost$pair_id == pair, ]
      cat("Pair:", pair, "| RF Freq:", rf_info$Freq, "| BOOST p:", boost_info$p_value, "\n")
    }
  }

  return(common_pairs_lenient)
}

lenient_results <- lenient_comparison()
#> === LENIENT COMPARISON (Top 100 each) ===
#Top 100 RF pairs overlap with top 100 BOOST pairs: 0

```

```{R, echo=FALSE, eval=FALSE}
# Extract all unique SNPs from RF interaction pairs
extract_rf_snps <- function(interaction_snps_counts) {

  # Split the interaction pairs and collect all SNP names
  all_snps <- c()

  for(i in 1:nrow(interaction_snps_counts)) {
    pair <- interaction_snps_counts$interaction_snps[i]
    snp_pair <- strsplit(as.character(pair), " ")[[1]]
    all_snps <- c(all_snps, snp_pair[1], snp_pair[2])
  }

  # Get unique SNPs
  unique_snps <- unique(all_snps)

  cat("RF interaction analysis:\n")
  cat("Total interaction pairs:", nrow(interaction_snps_counts), "\n")
  cat("Total SNP mentions:", length(all_snps), "\n") 
  cat("Unique SNPs involved:", length(unique_snps), "\n")

  return(unique_snps)
}

# Extract the SNPs
rf_snps <- extract_rf_snps(interaction_snps_counts[interaction_snps_counts$Freq > 1,]) #533 SNPS

# Get their indices in the original data
snp_names <- colnames(dt[,-1])
rf_snp_indices <- which(snp_names %in% rf_snps)

cat("SNP indices for RF-identified SNPs:\n")
cat("Number of indices found:", length(rf_snp_indices), "\n")
cat("Sample indices:", head(rf_snp_indices, 10), "\n")

# Boolean screening focused on RF-identified SNPs only
boolean_rf_focused <- function(snp_data, phenotype, rf_snp_indices, n_cores = 30) {

  n_rf_snps <- length(rf_snp_indices)
  total_rf_pairs <- choose(n_rf_snps, 2)

  cat("Focused Boolean screening on RF SNPs:\n")
  cat("RF-identified SNPs:", n_rf_snps, "\n")
  cat("Possible pairs among RF SNPs:", total_rf_pairs, "\n")
  cat("Using", n_cores, "cores\n")

  # Generate ALL possible pairs among RF SNPs (since the number is manageable)
  rf_pairs <- t(combn(rf_snp_indices, 2))

  cat("Testing all", nrow(rf_pairs), "pairs among RF SNPs\n")

  # Split into chunks for parallel processing
  chunk_size <- ceiling(nrow(rf_pairs) / n_cores)
  pair_chunks <- split(1:nrow(rf_pairs), ceiling(seq_along(1:nrow(rf_pairs)) / chunk_size))

  # Process in parallel using mclapply (shared memory)
  library(parallel)

  results_list <- mclapply(pair_chunks, function(chunk_indices) {
    chunk_results <- data.frame()

    for(idx in chunk_indices) {
      i <- rf_pairs[idx, 1]  # These are already the correct indices
      j <- rf_pairs[idx, 2]

      snp1 <- snp_data[, i]
      snp2 <- snp_data[, j]

      # Skip if no variation
      if (length(unique(snp1)) < 2 || length(unique(snp2)) < 2) next

      tryCatch({
        # Boolean operations
        and_op <- (snp1 >= 1) & (snp2 >= 1)
        or_op <- (snp1 >= 1) | (snp2 >= 1)
        xor_op <- (snp1 >= 1) != (snp2 >= 1)

        # Test association with phenotype
        and_test <- fisher.test(table(and_op, phenotype))
        or_test <- fisher.test(table(or_op, phenotype))
        xor_test <- fisher.test(table(xor_op, phenotype))

        chunk_results <- rbind(chunk_results, data.frame(
          SNP1_idx = i,
          SNP2_idx = j,
          SNP1_name = colnames(snp_data)[i],
          SNP2_name = colnames(snp_data)[j],
          boolean_and_p = and_test$p.value,
          boolean_or_p = or_test$p.value,
          boolean_xor_p = xor_test$p.value,
          best_boolean_p = min(and_test$p.value, or_test$p.value, xor_test$p.value),
          and_odds_ratio = as.numeric(and_test$estimate),
          or_odds_ratio = as.numeric(or_test$estimate),
          xor_odds_ratio = as.numeric(xor_test$estimate)
        ))

      }, error = function(e) {
        # Skip problematic pairs
      })
    }

    return(chunk_results)
  }, mc.cores = n_cores)

  # Combine results
  final_results <- do.call(rbind, results_list)

  if(nrow(final_results) > 0) {
    # Add FDR correction
    final_results$fdr_p_value <- p.adjust(final_results$best_boolean_p, method = "fdr")

    # Add pair_id
    final_results$pair_id <- apply(final_results[, c("SNP1_name", "SNP2_name")], 1,
                                  function(x) paste(sort(x), collapse = " "))

    # Sort by p-value
    final_results <- final_results[order(final_results$best_boolean_p), ]
  }

  cat("Completed! Tested", nrow(final_results), "pairs\n")
  return(final_results)
}

# Run the focused Boolean screening
cat("Starting focused Boolean screening on RF-identified SNPs...\n")

boost_time_focused <- system.time({
  boost_rf_focused <- boolean_rf_focused(
    dt[,-1],              # SNP data
    dt[,1],               # phenotype  
    rf_snp_indices,       # Indices of RF-identified SNPs
    n_cores = 60
  )
})

cat("Focused Boolean screening completed in", boost_time_focused[3], "seconds\n")
#Focused Boolean screening completed in 47.242 seconds
cat("Results summary:\n")
cat("Total pairs tested:", nrow(boost_rf_focused), "\n") #Total pairs tested: 141778 
cat("Significant at p < 0.05:", sum(boost_rf_focused$best_boolean_p < 0.05), "\n") #Significant at p < 0.05: 61106 
cat("Significant at FDR < 0.05:", sum(boost_rf_focused$fdr_p_value < 0.05), "\n")  #Significant at FDR < 0.05: 41651

print(head(boost_rf_focused, 10))

# Now we have a perfect comparison - same SNPs, different methods
direct_comparison <- function(rf_results, boost_focused_results) {

  # Find overlapping pairs
  common_pairs <- intersect(rf_results$interaction_snps, boost_focused_results$pair_id)

  cat("=== DIRECT COMPARISON (Same SNPs) ===\n")
  cat("RF pairs:", nrow(rf_results), "\n")
  cat("BOOST pairs tested:", nrow(boost_focused_results), "\n")
  cat("BOOST significant (FDR < 0.05):", sum(boost_focused_results$fdr_p_value < 0.05), "\n")
  cat("Overlapping pairs:", length(common_pairs), "\n")
  cat("Overlap percentage:", round(length(common_pairs) / nrow(rf_results) * 100, 2), "%\n")

  # Detailed analysis of overlapping pairs
  if(length(common_pairs) > 0) {
    overlap_details <- data.frame()

    for(pair in common_pairs) {
      rf_row <- rf_results[rf_results$interaction_snps == pair, ]
      boost_row <- boost_focused_results[boost_focused_results$pair_id == pair, ]

      overlap_details <- rbind(overlap_details, data.frame(
        pair_id = pair,
        rf_frequency = rf_row$Freq[1],
        boost_p_value = boost_row$best_boolean_p[1],
        boost_fdr_p_value = boost_row$fdr_p_value[1],
        boost_odds_ratio = boost_row$and_odds_ratio[1]  # or whichever was best
      ))
    }

    print("Overlapping pairs:")
    print(overlap_details[order(overlap_details$rf_frequency, decreasing = TRUE), ])

    # Correlation analysis
    if(nrow(overlap_details) > 1) {
      correlation <- cor(overlap_details$rf_frequency, -log10(overlap_details$boost_p_value))
      cat("Correlation between RF frequency and BOOST -log10(p):", round(correlation, 3), "\n")
    }
  }

  return(list(
    common_pairs = common_pairs,
    overlap_details = if(exists("overlap_details")) overlap_details else data.frame()
  ))
}

# Perform direct comparison
direct_results <- direct_comparison(interaction_snps_counts[interaction_snps_counts$Freq > 1,], boost_rf_focused)
#=== DIRECT COMPARISON (Same SNPs) ===
#RF pairs: 640 
#BOOST pairs tested: 141778 
#BOOST significant (FDR < 0.05): 41651 
#Overlapping pairs: 640 
#Overlap percentage: 100 %

head(direct_results$overlap_details)

# Save all results
#save(rf_snps, rf_snp_indices, boost_rf_focused, direct_results, 
#     file = "./script3/focused_rf_boost_comparison.Rdata")
#load("./script3/focused_rf_boost_comparison.Rdata")

```

```{R, echo=FALSE, eval=FALSE}
# Analyze the relationship between RF and BOOST findings
analyze_rf_boost_relationship <- function(direct_results) {

  overlap_data <- direct_results$overlap_details

  cat("=== RF vs BOOST RELATIONSHIP ANALYSIS ===\n")

  # Summary statistics
  cat("RF frequency range:", range(overlap_data$rf_frequency), "\n")
  cat("BOOST p-value range:", range(overlap_data$boost_p_value), "\n")
  cat("Correlation (RF freq vs -log10(p)):", round(cor(overlap_data$rf_frequency, -log10(overlap_data$boost_p_value)), 3), "\n")

  # Categorize RF pairs by frequency
  high_freq_rf <- overlap_data[overlap_data$rf_frequency >= 50, ]
  med_freq_rf <- overlap_data[overlap_data$rf_frequency >= 10 & overlap_data$rf_frequency < 50, ]
  low_freq_rf <- overlap_data[overlap_data$rf_frequency < 10, ]

  cat("\n=== FREQUENCY CATEGORIES ===\n")
  cat("High frequency RF pairs (≥50):", nrow(high_freq_rf), "\n")
  cat("- Mean BOOST -log10(p):", round(mean(-log10(high_freq_rf$boost_p_value)), 2), "\n")
  cat("- BOOST FDR < 0.05:", sum(high_freq_rf$boost_fdr_p_value < 0.05), "/", nrow(high_freq_rf), "\n")

  cat("Medium frequency RF pairs (10-49):", nrow(med_freq_rf), "\n")
  cat("- Mean BOOST -log10(p):", round(mean(-log10(med_freq_rf$boost_p_value)), 2), "\n")
  cat("- BOOST FDR < 0.05:", sum(med_freq_rf$boost_fdr_p_value < 0.05), "/", nrow(med_freq_rf), "\n")

  cat("Low frequency RF pairs (2-9):", nrow(low_freq_rf), "\n")
  cat("- Mean BOOST -log10(p):", round(mean(-log10(low_freq_rf$boost_p_value)), 2), "\n")
  cat("- BOOST FDR < 0.05:", sum(low_freq_rf$boost_fdr_p_value < 0.05), "/", nrow(low_freq_rf), "\n")

  return(list(high = high_freq_rf, medium = med_freq_rf, low = low_freq_rf))
}

# Run the analysis
freq_analysis <- analyze_rf_boost_relationship(direct_results)
## === RF vs BOOST RELATIONSHIP ANALYSIS ===
## RF frequency range: 2 466 
## BOOST p-value range: 5.977923e-71 0.8822303 
## Correlation (RF freq vs -log10(p)): 0.149

## === FREQUENCY CATEGORIES ===
## High frequency RF pairs (≥50): 14 
## - Mean BOOST -log10(p): 33.3 
## - BOOST FDR < 0.05: 12 / 14 
## Medium frequency RF pairs (10-49): 31 
## - Mean BOOST -log10(p): 28.6 
## - BOOST FDR < 0.05: 30 / 31 
## Low frequency RF pairs (2-9): 595 
## - Mean BOOST -log10(p): 24.77 
## - BOOST FDR < 0.05: 537 / 595

# Compare what each method considers "significant"
method_comparison_analysis <- function(rf_data, boost_data) {

  cat("=== METHOD SENSITIVITY COMPARISON ===\n")

  # RF considers Freq > 1 as significant (your filter)
  rf_significant <- nrow(rf_data)  # All 640 pairs

  # BOOST considers FDR < 0.05 as significant  
  boost_significant <- sum(boost_data$fdr_p_value < 0.05)
  boost_very_significant <- sum(boost_data$fdr_p_value < 0.01)

  cat("RF significant pairs (Freq > 1):", rf_significant, "\n")
  cat("BOOST significant pairs (FDR < 0.05):", boost_significant, "\n")
  cat("BOOST very significant pairs (FDR < 0.01):", boost_very_significant, "\n")

  # How many RF pairs are also BOOST significant?
  rf_pairs_also_boost_sig <- sum(direct_results$overlap_details$boost_fdr_p_value < 0.05)
  rf_pairs_also_boost_very_sig <- sum(direct_results$overlap_details$boost_fdr_p_value < 0.01)

  cat("\nOverlap analysis:\n")
  cat("RF pairs also BOOST significant (FDR < 0.05):", rf_pairs_also_boost_sig, "/", rf_significant, 
      "(", round(rf_pairs_also_boost_sig/rf_significant*100, 1), "%)\n")
  cat("RF pairs also BOOST very significant (FDR < 0.01):", rf_pairs_also_boost_very_sig, "/", rf_significant,
      "(", round(rf_pairs_also_boost_very_sig/rf_significant*100, 1), "%)\n")

  # What about BOOST-only significant pairs?
  boost_only_pairs <- boost_data[!(boost_data$pair_id %in% rf_data$interaction_snps) & 
                                 boost_data$fdr_p_value < 0.05, ]

  cat("BOOST-only significant pairs (not in RF Freq > 1):", nrow(boost_only_pairs), "\n")

  return(list(
    rf_sig = rf_significant,
    boost_sig = boost_significant, 
    overlap_sig = rf_pairs_also_boost_sig,
    boost_only = boost_only_pairs
  ))
}

# Run method comparison
method_comp <- method_comparison_analysis(
  interaction_snps_counts[interaction_snps_counts$Freq > 1,], 
  boost_rf_focused
)
## === METHOD SENSITIVITY COMPARISON ===
## RF significant pairs (Freq > 1): 640 
## BOOST significant pairs (FDR < 0.05): 41651 
## BOOST very significant pairs (FDR < 0.01): 24334

## Overlap analysis:
## RF pairs also BOOST significant (FDR < 0.05): 579 / 640 ( 90.5 %)
## RF pairs also BOOST very significant (FDR < 0.01): 572 / 640 ( 89.4 %)
## BOOST-only significant pairs (not in RF Freq > 1): 41072

# Summary table 
create_final_summary <- function() {

  summary_stats <- data.frame(
    Metric = c(
      "Total unique SNPs tested",
      "Total possible pairs", 
      "RF significant pairs (Freq > 1)",
      "BOOST pairs tested",
      "BOOST significant (FDR < 0.05)",
      "RF-BOOST overlap",
      "Correlation (RF freq vs BOOST -log10p)"
    ),
    Value = c(
      "533",
      format(choose(533, 2), big.mark = ","),
      "640", 
      format(nrow(boost_rf_focused), big.mark = ","),
      format(sum(boost_rf_focused$fdr_p_value < 0.05), big.mark = ","),
      "640 (100%)",
      "0.149"
    )
  )

  print(summary_stats)

  cat("\n=== KEY FINDINGS FOR PAPER ===\n")
  cat("1. When testing the same SNPs, both methods detect interactions\n")
  cat("2. RF frequency correlates weakly with BOOST significance (r=0.149)\n") 
  cat("3. RF identifies", nrow(direct_results$overlap_details), "pairs based on co-occurrence\n")
  cat("4. BOOST identifies", sum(boost_rf_focused$fdr_p_value < 0.05), "pairs based on Boolean logic\n")
  cat("5. Methods capture different aspects of epistatic interactions\n")
}

create_final_summary()
create_final_summary()
#                                  Metric      Value
#1               Total unique SNPs tested        533
#2                   Total possible pairs    141,778
#3        RF significant pairs (Freq > 1)        640
#4                     BOOST pairs tested    141,778
#5         BOOST significant (FDR < 0.05)     41,651
#6                       RF-BOOST overlap 640 (100%)
#7 Correlation (RF freq vs BOOST -log10p)      0.149

#=== KEY FINDINGS FOR PAPER ===
#1. When testing the same SNPs, both methods detect interactions
#2. RF frequency correlates weakly with BOOST significance (r=0.149)
#3. RF identifies 640 pairs based on co-occurrence
#4. BOOST identifies 41651 pairs based on Boolean logic
#5. Methods capture different aspects of epistatic interactions

# Save final results
#save(direct_results, freq_analysis, method_comp, 
#     file = "./script3/final_rf_boost_comparison.Rdata")
#load( file = "./script3/final_rf_boost_comparison.Rdata")

```

- things to do for comparison with BOOST (first in the Supplementary and then in the paper)

   -   Methods section - Clean description of both RF and Boolean screening approaches
   -  Results section - Key findings table and interpretation
   -  Comparison figures - Correlation plots, overlap analysis
   -  Discussion points - Why methods detect different interaction types